---
title: "Case Study"
author: "Team 3"
format: revealjs
editor: visual
---

```{r data}
# first, import all necessary packages
library(ggplot2)
library(dplyr) # seems very useful, ill use it just for conditional removal of rows from our data

# get our data set
df <- `2026.08.MD`

# take the dataset relevant to us
ds <- df[c("V1","V7","V105","V77")] 

# rename columns to be more expressive
colnames(ds) <- c("Date","IndustrialProduction","ConsumerPriceIndex","FedFundRate") 

# remove the weird two rows from the original data
ds <- ds[-c(1:2), ] 

# check types of the columns, they are all charater type
print("Check the types of the columns to see if we must adjust them")
sapply(ds, class) 

# they were all character type, convert to the correct respective types
#as.numeric(ds$`Industrial Production`)
#as.numeric(ds$`Consumer Price Index`)
#as.numeric(ds$`Fed Fund Rate`)
#as.Date(ds$Date, format="%d/%m/%Y")

ds <- transform(
  ds, `Date` = as.Date(ds$Date, format="%d/%m/%Y"), `IndustrialProduction` = as.numeric(ds$`IndustrialProduction`), `ConsumerPriceIndex` = as.numeric(ds$`ConsumerPriceIndex`), `FedFundRate` = as.numeric(ds$`FedFundRate`))

# check they are the correct type
print("Check the types of the columns again to ensure they are correct")
sapply(ds, class) 
ds
# we now have the correct types

# now we will remove data from the end of 2019 onwards to avoid looking at the anomalous data from covid-19. we use dplyr
ds <- ds %>%
  filter(`Date`<"2020-01-01")

ds
```

Now that we have cleaned and prepared our data, we can move onto to exploring it.

```{r visual}
# get an initial look into them by plotting them
ggplot(aes(x = `Date`, y = `IndustrialProduction`), data = ds) + geom_point()
ggplot(aes(x = `Date`, y = `ConsumerPriceIndex`), data = ds) + geom_point()
ggplot(aes(x = `Date`, y = `FedFundRate`), data = ds) + geom_point()
```

Apply the necessary transformations for the time series according to the Appendix of McCracken and Ng (2016):

i.) Δ*xt* for FedFundRate;\
ii.) Δ\^2log (*xt*) for ConsumerPriceIndex;\
iii.) Δlog (*xt*) for IndustrialProduction.

```{r}
# apply the necessary transformations to make the series stationary

# FedFundRate (FFR)
d_FFR <- diff(ds$FedFundRate)

# ConsumerPriceIndex (CPI)
d2_log_cpi <- diff(log(ds$ConsumerPriceIndex), differences = 2) # differencing two times

# IndustrialProduction (IP)
d_log_IP <- diff(log(ds$IndustrialProduction))

# Transform the dataframe. Add new colums for stationary series
ds_transformed <- ds %>%
  mutate(
    d_log_IP   = c(NA, diff(log(IndustrialProduction))), # Fill first row with NA. Differencing deletes first observation.
    d2_log_cpi = c(NA, NA, diff(log(ConsumerPriceIndex), differences = 2)), # Fill first two rows with NA to make up for the differencing.
    d_FFR  = c(NA, diff(FedFundRate)) # Fill first row with NA to make up for differencing
  ) %>%
  
# Drop the first 2 rows where differencing CPI was filled with NAs to maintain a balanced panel. Dataframe begins at row 3.
filter(!is.na(d2_log_cpi))
  
# Plot the transformed series
plot(ds_transformed$Date, ds_transformed$d_log_IP)
plot(ds_transformed$Date, ds_transformed$d2_log_cpi)
plot(ds_transformed$Date, ds_transformed$d_FFR)
```

The plots show the transformed time series. The intuition behind the time series is now:

i.) d_FFR –\> monthly Fed Fund Rate adjustment;\
ii.) d2_log_cpi –\> acceleration of CPI (analogous to acceleration being the second derivative of distance);\
iii.) d_log_IP –\> Month-over-month IP growth rate.
