---
title: "Case 2: Structural Vector AutoRegressive Models, Local Projections and Macro-economic Policy Analysis"
subtitle: "EBC4008 -- Time Series Analytics"
author:
  - "Can Idil"
  - "Sem Huizinga"
  - "Juan Diego Gaviño"
format:
  beamer:
    slide-level: 2
    header-includes:
      - |
        \definecolor{navyblue}{RGB}{10,32,72}

        \setbeamertemplate{navigation symbols}{}

        % Main presentation colors
        \setbeamercolor{normal text}{fg=navyblue,bg=white}
        \setbeamercolor{frametitle}{fg=navyblue}
        \setbeamercolor{structure}{fg=navyblue}

        % Header/footer colors
        \setbeamercolor{headlinebar}{bg=navyblue,fg=white}
        \setbeamercolor{footlinebar}{bg=navyblue,fg=white}

        % Header
        \setbeamertemplate{headline}{
          \leavevmode
          \begin{beamercolorbox}[
            wd=\paperwidth,
            ht=3.5ex,
            dp=1.0ex
          ]{headlinebar}

            \hbox to \paperwidth{%

              % Left: current section
              \rlap{%
                \hspace{1em}
                \scriptsize \insertsectionhead
              }%

              \hfill

              % Right
              \llap{%
                \scriptsize Case 2
                \hspace{1em}
              }%
            }

          \end{beamercolorbox}
        }

        % Footer
        \setbeamertemplate{footline}{
          \leavevmode
          \begin{beamercolorbox}[
            wd=\paperwidth,
            ht=2.8ex,
            dp=1.0ex
          ]{footlinebar}

            \hbox to \paperwidth{%

              % Left
              \rlap{%
                \hspace{1em}
                \tiny Maastricht University
              }%

              \hfill

              % Exact center
              \tiny EBC4008 -- Time Series Analytics

              \hfill

              % Right
              \llap{%
                \tiny \insertframenumber{} / \inserttotalframenumber
                \hspace{1em}
              }%
            }

          \end{beamercolorbox}
        }
editor: visual
execute:
  echo: true
  warning: false
  message: false
---

## Table of Contents {.t}

\normalsize

\begin{enumerate}
  \item Structural VARs
  \item Impulse Response Functions for recursive SVARs
  \item VAR/LP IRFs Comparison
  \item Structural IRFs using LP
  \item Shock Proxies
\end{enumerate}





```{r}
#| include: false

library(vars)
library(lpirfs)

set.seed(123)

# Data preparation
raw <- read.csv("v2-2026-rev-08-MD.csv", check.names = FALSE)

dat <- raw[, c("sasdate", "INDPRO", "CPIAUCSL", "FEDFUNDS")]
names(dat)[1] <- "Date"

dat$Date <- as.Date(dat$Date, format = "%m/%d/%Y")
dat <- dat[!is.na(dat$Date) & dat$Date <= as.Date("2019-12-31"), ]
dat <- dat[order(dat$Date), ]

# Transformations
dat_transformed <- data.frame(
  Date = dat$Date,
  INDPRO = c(NA, diff(log(dat$INDPRO))),
  CPIAUCSL = c(NA, NA, diff(log(dat$CPIAUCSL), differences = 2)),
  FEDFUNDS = c(NA, diff(dat$FEDFUNDS))
)

dat_transformed <- dat_transformed[-c(1, 2), ]

Y <- data.frame(
  Date = dat_transformed$Date,
  IP = dat_transformed$INDPRO,
  CPI = dat_transformed$CPIAUCSL,
  FFR = dat_transformed$FEDFUNDS
)

variables <- c("IP", "CPI", "FFR")
x <- as.matrix(Y[, variables])

stopifnot(all(is.finite(x)))

# Reduced-form VAR
p_aic <- 14
fit_aic <- VAR(x, p = p_aic, type = "const")

irf_horizon <- 48

irf_reduced <- irf(
  fit_aic,
  impulse = variables,
  response = variables,
  n.ahead = irf_horizon,
  ortho = FALSE,
  boot = FALSE
)

# IRFs in levels
to_levels <- function(irf_matrix) {
  data.frame(
    Horizon = 0:(nrow(irf_matrix) - 1),
    IP = cumsum(irf_matrix[, "IP"]),
    CPI = cumsum(cumsum(irf_matrix[, "CPI"])),
    FFR = cumsum(irf_matrix[, "FFR"])
  )
}

policy_reduced_levels <- to_levels(
  irf_reduced$irf$FFR[, variables, drop = FALSE]
)

# IRF comparison plots
plot_comparison <- function(data_list, labels,
                            line_types = seq_along(data_list)) {

  responses <- c("IP", "CPI", "FFR")

  oldpar <- par(no.readonly = TRUE)
  on.exit({
    layout(1)
    par(oldpar)
  })

  layout(
    matrix(c(1, 2, 3, 4, 4, 4), nrow = 2, byrow = TRUE),
    heights = c(5, 0.65)
  )

  par(
    mar = c(2.8, 3.0, 1.8, 0.5),
    oma = c(1.5, 0, 0, 0),
    pty = "s",
    cex = 0.8
  )

  for (j in seq_along(responses)) {

    response <- responses[j]
    horizon <- data_list[[1]]$Horizon

    all_y <- unlist(lapply(
      data_list,
      function(z) z[[response]]
    ))

    plot(
      horizon,
      data_list[[1]][[response]],
      type = "l",
      ylim = range(c(all_y, 0)),
      xlab = "",
      ylab = "",
      main = response,
      axes = FALSE,
      lwd = 1.4,
      lty = line_types[1]
    )

    if (length(data_list) > 1) {
      for (i in 2:length(data_list)) {
        lines(
          data_list[[i]]$Horizon,
          data_list[[i]][[response]],
          lwd = 1.4,
          lty = line_types[i]
        )
      }
    }

    abline(h = 0, lty = 3, col = "grey40")

    axis(
      1,
      at = seq(0, irf_horizon, by = 12),
      cex.axis = 0.7
    )

    axis(2, cex.axis = 0.7)
  }

  # Shared legend
  par(mar = c(0, 0, 0, 0), pty = "m")
  plot.new()

  legend(
    "center",
    legend = labels,
    lty = line_types,
    lwd = 1.4,
    horiz = TRUE,
    bty = "n",
    cex = 0.85
  )

  mtext(
    "Horizon (months)",
    side = 1,
    outer = TRUE,
    line = 0.2,
    cex = 0.8
  )
}
```



## Recap -- Part I {.t}

\normalsize

\textcolor{navyblue}{$\bullet$} Reduced-form VAR(14) was estimated for

\footnotesize
$$
\Delta\log(INDPRO_t),\qquad
\Delta^2\log(CPI_t),\qquad
\Delta FFR_t
$$

\vspace{0.35em}

\normalsize
\textcolor{navyblue}{$\bullet$} Reduced-form IRFs were computed over 48 months with bootstrap confidence intervals

\vspace{0.35em}

\textcolor{navyblue}{$\bullet$} The responses were accumulated to recover the variables in levels

\vspace{0.35em}

\textcolor{navyblue}{$\bullet$} A positive FFR innovation is followed by an increase in the CPI level, producing a price-puzzle-like pattern

\vspace{0.35em}

\centering

\normalsize
\textcolor{navyblue}{$\Rightarrow$} \textit{FFR innovation is \textbf{not yet structurally identified}, motivating the \textbf{SVAR} analysis}


# Structural VARs

## Reduced-Form VAR versus Structural VAR {.t}

\vspace{0.5em}

\normalsize

\textcolor{navyblue}{$\bullet$} Reduced-form VAR: statistical dynamics and forecast errors

\vspace{-0.8em}

\footnotesize
$$
y_t=A_1y_{t-1}+\cdots+A_py_{t-p}+u_t
$$

\vspace{0.15em}

\normalsize

\textcolor{navyblue}{$\bullet$} Structural VAR: contemporaneous relationships and economic shocks

\vspace{-1.5em}

\footnotesize
$$
B_0y_t=B_1y_{t-1}+\cdots+B_py_{t-p}+w_t
$$

\vspace{-0.3em}

\scriptsize
$$
A_i=B_0^{-1}B_i,
\qquad
u_t=B_0^{-1}w_t,
\qquad
E(w_tw_t')=I_K
$$

\vspace{0.3em}

\footnotesize

\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
$u_t$ may combine several underlying economic disturbances

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
$B_0$ captures contemporaneous relationships; $w_t$ is the vector of structural shocks

\end{itemize}

\footnotesize
\underline{Objective:} Recover economically interpretable shocks from the reduced-form errors

---


## From Reduced-Form to Structural VAR {.t}

\vspace{0.8em}

\normalsize

\textcolor{navyblue}{$\bullet$} Premultiplying the reduced-form companion VAR by $\mathcal B_0$ yields the structural representation

\vspace{0.3em}

\small
$$
\mathcal B_0Y_t
=
\underbrace{\mathcal B_0 A}_{\mathcal B_1}
Y_{t-1}
+
\underbrace{\mathcal B_0 U_t}_{W_t}
$$

\vspace{0.3em}



\small where \quad \scriptsize $Y_t,U_t,W_t\in\mathbb R^{Kp}$, $A,\mathcal B_0,\mathcal B_1\in\mathbb R^{Kp\times Kp}$, and $\mathcal B_0=\operatorname{diag}(B_0,I_K,\ldots,I_K)$.

\vspace{0.8em}

\normalsize

\textcolor{navyblue}{$\bullet$} Both representations describe the same dynamics, but the structural interpretation depends on identifying $B_0$


## Identification Problem {.t}

\vspace{0.5em}

\normalsize

\textcolor{navyblue}{$\bullet$} The reduced-form covariance matrix provides information about $B_0^{-1}$

\vspace{-1.5em}

\footnotesize
$$
\Sigma_u=B_0^{-1}(B_0^{-1})'
$$

\vspace{0.25em}

\normalsize

\textcolor{navyblue}{$\bullet$} However, it provides fewer independent equations than unknown coefficients

\vspace{-0.7em}

\footnotesize
$$
\underbrace{K^2}_{\text{Unknowns}}
-
\underbrace{\frac{K(K+1)}{2}}_{\text{Equations}}
=
\underbrace{\frac{K(K-1)}{2}}_{\text{Restrictions}}
$$

\vspace{0.25em}

\normalsize

\textcolor{navyblue}{$\bullet$} For the trivariate VAR:

\vspace{-0.7em}

\footnotesize
$$
9-6=\boxed{3\text{ restrictions}}
$$

\vspace{0.35em}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} At least three independent restrictions are needed; the rank condition must also hold


---

## Types of Identification Restrictions {.t}

\vspace{0.5em}

\normalsize

\textcolor{navyblue}{$\bullet$} Structural restrictions can constrain the effects of shocks at different horizons

\vspace{0.2em}

\footnotesize
\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
\textbf{Short-run:} zero, equality or proportionality restrictions on contemporaneous coefficients

\vspace{0.15em}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
\textbf{Long-run:} restrictions on cumulative or permanent effects of structural shocks

\vspace{0.15em}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
\textbf{Sign:} restrictions on the direction of responses over selected horizons

\end{itemize}

\vspace{0.3em}

\normalsize

\textcolor{navyblue}{$\bullet$} Economic justification may arise from theory, information delays, physical constraints or institutional knowledge

\vspace{0.3em}

\normalsize

\textcolor{navyblue}{$\bullet$} Here: Short-run exclusion restrictions to identify monetary-policy shocks



# Impulse Response Functions for Recursive SVARs

```{r}
#| include: false

# Recursive identification
Bmat_recursive <- matrix(NA_real_, 3, 3)
Bmat_recursive[upper.tri(Bmat_recursive)] <- 0

# Baseline: IP, CPI, FFR
svar_recursive <- SVAR(
  fit_aic,
  Bmat = Bmat_recursive,
  lrtest = FALSE
)

irf_svar <- irf(
  svar_recursive,
  impulse = "FFR",
  response = variables,
  n.ahead = irf_horizon,
  boot = FALSE
)

# Alternative: CPI, IP, FFR
order_swap <- c("CPI", "IP", "FFR")

fit_swap <- VAR(
  x[, order_swap, drop = FALSE],
  p = p_aic,
  type = "const"
)

svar_swap <- SVAR(
  fit_swap,
  Bmat = Bmat_recursive,
  lrtest = FALSE
)

irf_swap <- irf(
  svar_swap,
  impulse = "FFR",
  response = variables,
  n.ahead = irf_horizon,
  boot = FALSE
)

# Alternative: FFR, IP, CPI
order_first <- c("FFR", "IP", "CPI")

fit_first <- VAR(
  x[, order_first, drop = FALSE],
  p = p_aic,
  type = "const"
)

svar_first <- SVAR(
  fit_first,
  Bmat = Bmat_recursive,
  lrtest = FALSE
)

irf_first <- irf(
  svar_first,
  impulse = "FFR",
  response = variables,
  n.ahead = irf_horizon,
  boot = FALSE
)

# Orient monetary-policy shocks
orient_policy <- function(irf_matrix) {
  if (irf_matrix[1, "FFR"] < 0) {
    irf_matrix <- -irf_matrix
  }
  irf_matrix
}

# Responses in levels
policy_svar_levels <- to_levels(
  orient_policy(irf_svar$irf$FFR[, variables, drop = FALSE])
)

policy_swap_levels <- to_levels(
  orient_policy(irf_swap$irf$FFR[, variables, drop = FALSE])
)

policy_first_levels <- to_levels(
  orient_policy(irf_first$irf$FFR[, variables, drop = FALSE])
)
```




## Recursive Identification through Ordering {.t}

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Recursive identification imposes a lower-triangular structural impact matrix

\vspace{-0.6em}

\footnotesize
$$
u_t=
\underbrace{
\begin{bmatrix}
b_{11}^{0}&0&0\\
b_{21}^{0}&b_{22}^{0}&0\\
b_{31}^{0}&b_{32}^{0}&b_{33}^{0}
\end{bmatrix}
}_{B_0^{-1}}
w_t
$$

\vspace{0.2em}

\footnotesize

\quad\textcolor{navyblue}{\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}} The three zeros satisfy the order condition for $K=3$; the Cholesky decomposition selects the corresponding lower-triangular impact matrix

\vspace{0.35em}

\normalsize

\textcolor{navyblue}{$\bullet$} Structural impulse responses are obtained from the reduced-form responses

\vspace{-1.3em}

\footnotesize
$$
\Theta_h=\Phi_hB_0^{-1},
\qquad h=0,\ldots,H
$$

\vspace{0.25em}

\normalsize

\textcolor{navyblue}{$\bullet$} The ordering determines which shocks can affect each variable contemporaneously

---


## Ordering the Monetary-Policy VAR {.t}

\vspace{0.55em}

\normalsize

\textcolor{navyblue}{$\bullet$} We adopt the ordering

\vspace{-0.65em}

\footnotesize
$$
y_t=(IP_t,\ CPI_t,\ FFR_t)'
$$

\vspace{0.3em}

\normalsize

\textcolor{navyblue}{$\bullet$} Economic motivation:

\vspace{0.15em}

\footnotesize
\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
\textbf{IP and CPI:} assumed not to respond within the month to unexpected monetary-policy changes

\vspace{0.2em}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
\textbf{FFR:} allowed to react contemporaneously to real activity and price developments

\end{itemize}

\vspace{0.35em}

\normalsize

\textcolor{navyblue}{$\bullet$} The last innovation is interpreted as the monetary-policy shock

\vspace{-0.7em}

\footnotesize
$$
w_{3t}\equiv w_t^{MP}
$$

\vspace{0.25em}

\small

\quad\textcolor{navyblue}{$\Rightarrow$} Only the policy shock is assigned an economic interpretation (semistructural SVAR)

---


## Estimating Recursive SVAR IRFs {.t}

\vspace{0.7em}

\normalsize

\textcolor{navyblue}{$\bullet$} We use \texttt{SVAR()} with recursive zero restrictions on the structural impact matrix

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} \texttt{irf()} computes structural responses from the VAR(14) over 48 months

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Structural shocks have unit variance and are oriented to raise FFR on impact

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Responses are accumulated to recover levels

\vspace{0.55em}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} The figures compare point estimates across identification assumptions

---


## Structural Monetary-Policy IRFs

```{r}
#| echo: false
#| fig-width: 9.0
#| fig-height: 3.7


plot_comparison(
  list(
    policy_reduced_levels,
    policy_svar_levels
  ),
  labels = c(
    "Reduced-form VAR",
    "SVAR"
  ),
  line_types = c(4, 1)
)
```

---

## Structural Monetary-Policy IRFs {.t}

\vspace{0.65em}

\normalsize

\textcolor{navyblue}{$\bullet$} With FFR ordered last, structural and reduced-form policy IRFs are proportional

\vspace{-0.7em}

\footnotesize
$$
\Theta_h e_3=b_{33}^{0}\Phi_h e_3
$$

\vspace{0.4em}

\normalsize

\textcolor{navyblue}{$\bullet$} Both curves have identical profiles; their magnitudes differ through shock normalization

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} IP declines, while the CPI level rises following a positive FFR shock

\vspace{0.55em}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} The price-puzzle pattern persists under recursive identification

---


## Ordering Sensitivity


```{r}
#| echo: false
#| fig-width: 9.0
#| fig-height: 3.7

plot_comparison(
  list(
    policy_svar_levels,
    policy_swap_levels,
    policy_first_levels
  ),
  labels = c(
    "IP, CPI, FFR",
    "CPI, IP, FFR",
    "FFR, IP, CPI"
  ),
  line_types = c(1, 3, 5)
)
```

---

## Ordering Sensitivity {.t}

\vspace{0.65em}

\normalsize

\textcolor{navyblue}{$\bullet$} \textbf{FFR last:} Swapping IP and CPI leaves the monetary-policy IRFs unchanged

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} \textbf{FFR first:} IP and CPI can respond immediately to policy shocks, changing the estimated responses

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Unit-variance structural shocks can also produce different FFR impact magnitudes across orderings

\vspace{0.55em}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} Ordering comparisons reveal sensitivity, but may represent different causal experiments



# VAR/LP IRFs Comparison

```{r}
#| include: false

# Reduced-form local projections
lp_irfs <- array(
  NA_real_,
  dim = c(irf_horizon + 1, 3, 3),
  dimnames = list(NULL, variables, variables)
)

for (h in 0:irf_horizon) {

  idx <- (p_aic + 1):(nrow(x) - h)

  X_h <- cbind(
    1,
    x[idx, , drop = FALSE]
  )

  for (lag in 1:p_aic) {
    X_h <- cbind(
      X_h,
      x[idx - lag, , drop = FALSE]
    )
  }

  fit_h <- lm.fit(
    X_h,
    x[idx + h, , drop = FALSE]
  )

  lp_irfs[h + 1, , ] <- t(
    fit_h$coefficients[2:4, , drop = FALSE]
  )
}

# LP responses in levels
lp_all_levels <- setNames(
  lapply(variables, function(v) {
    to_levels(lp_irfs[, , v])
  }),
  variables
)

# VAR responses in levels
var_all_levels <- setNames(
  lapply(variables, function(v) {
    to_levels(
      irf_reduced$irf[[v]][, variables, drop = FALSE]
    )
  }),
  variables
)

# 3x3 VAR-LP comparison
plot_var_lp_grid <- function() {

  oldpar <- par(no.readonly = TRUE)
  on.exit({
    layout(1)
    par(oldpar)
  })

  layout(
    matrix(c(
      1, 2, 3,
      4, 5, 6,
      7, 8, 9,
      10, 10, 10
    ), nrow = 4, byrow = TRUE),
    heights = c(1, 1, 1, 0.35),
    respect = FALSE
  )

  par(
    mar = c(1.6, 2.4, 1.1, 0.3),
    oma = c(0.3, 3.0, 2.1, 0.2),
    pty = "m",
    cex = 0.8,
    mgp = c(1.3, 0.3, 0)
  )

  for (i in seq_along(variables)) {
    for (j in seq_along(variables)) {

      impulse <- variables[i]
      response <- variables[j]

      var_y <- var_all_levels[[impulse]][[response]]
      lp_y <- lp_all_levels[[impulse]][[response]]
      horizon <- 0:irf_horizon

      plot(
        horizon, var_y,
        type = "n",
        ylim = range(c(var_y, lp_y, 0)),
        xlab = "",
        ylab = "",
        axes = FALSE
      )

      abline(h = 0, lty = 3, col = "grey40")

      lines(horizon, var_y, lty = 1, lwd = 1.5)
      lines(horizon, lp_y, lty = 2, lwd = 1.5)

      if (i == 3) {
        axis(1, at = seq(0, 48, by = 12), cex.axis = 0.65)
      }

      axis(2, cex.axis = 0.6, las = 1)

      # Column labels
      if (i == 1) {
        mtext(
          response,
          side = 3,
          line = 0.2,
          font = 2,
          cex = 0.85
        )
      }

      # Row labels
      if (j == 1) {
        mtext(
          impulse,
          side = 2,
          line = 2.0,
          font = 2,
          cex = 0.85
        )
      }
    }
  }

  # Shared legend
  par(mar = c(0, 0, 0, 0), pty = "m")
  plot.new()

  legend(
    "top",
    legend = c("Reduced-form VAR", "Local Projections"),
    lty = c(1, 2),
    lwd = 1.5,
    horiz = TRUE,
    bty = "n",
    cex = 0.85
  )

  text(0.5, 0.12, "Horizon (months)", cex = 0.8)

  # Shared grid labels
  mtext(
    "Response",
    side = 3,
    outer = TRUE,
    line = 0.6,
    font = 2,
    cex = 0.85
  )

  mtext(
    "Impulse",
    side = 2,
    outer = TRUE,
    line = 1.5,
    font = 2,
    cex = 0.85
  )
}

```


## Local Projections {.t}

\vspace{0.6em}

\normalsize

\textcolor{navyblue}{$\bullet$} Local projections (Jordà, 2005) estimate IRFs directly through horizon-specific regressions

\vspace{-0.7em}

\footnotesize
$$
y_{t+h}
=
\alpha_h+\Psi_h y_t
+\sum_{\ell=1}^{p}\Gamma_{\ell,h}y_{t-\ell}
+v_{t+h}^{(h)}
$$

\vspace{0.3em}

\normalsize

\textcolor{navyblue}{$\bullet$} A separate regression is estimated for each horizon $h=0,\ldots,H$

\vspace{0.3em}

\normalsize

\textcolor{navyblue}{$\bullet$} The coefficients $\widehat{\Psi}_h$ give the reduced-form responses to unit innovations, holding the other contemporaneous innovations fixed

\vspace{0.35em}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} IRFs are obtained directly, without iterating a fitted VAR

---


## VARs versus Local Projections {.t}

\vspace{0.5em}

\centering
\footnotesize

\renewcommand{\arraystretch}{1.7}
\setlength{\tabcolsep}{5pt}

\begin{tabular}{p{1.8cm}p{3.8cm}p{3.8cm}}

\hline

& \textbf{VAR} & \textbf{Local Projections} \\

\hline

\textbf{Estimation}
& Estimates one dynamic system; IRFs are obtained by iterating its estimated coefficients
& Estimates a separate regression at each horizon, without imposing a common dynamic structure \\

\hline

\textbf{Advantages}
& Typically more precise and smoother IRFs, as information is shared across horizons
& More flexible response shapes; often less bias when the VAR lag structure is misspecified \\

\hline

\textbf{Disadvantages}
& Misspecified dynamics can introduce bias that propagates across horizons
& Higher sampling variance, especially at longer horizons; potentially irregular IRFs \\

\hline

\end{tabular}

\vspace{0.7em}

\raggedright
\footnotesize

\textcolor{navyblue}{$\Rightarrow$} Both target the same population IRFs with sufficiently rich lag structures, but differ in their finite-sample bias-variance trade-off


---


## Reduced-Form LP Implementation {.t}

\vspace{0.75em}

\normalsize

\textcolor{navyblue}{$\bullet$} Estimate horizon-specific OLS regressions using \texttt{lm.fit()}

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Same IP, CPI and FFR transformations, with 14 lagged controls and 48-month horizons

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Contemporaneous regressors enter jointly, yielding nine unorthogonalized reduced-form IRFs

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Accumulate responses once for IP and FFR, twice for CPI

\vspace{0.55em}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} VAR and LP estimates are compared for the same unit innovations

---


## Reduced-Form IRFs: VAR versus LP

```{r}
#| echo: false
#| fig-width: 9.0
#| fig-height: 5.5
#| out-width: 100%
#| fig-align: center

plot_var_lp_grid()
```

---

## Reduced-Form IRFs: VAR versus LP {.t}

\vspace{0.6em}

\normalsize

\textcolor{navyblue}{$\bullet$} All nine IRFs coincide on impact, with larger differences at longer horizons

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} LP estimates exhibit less smoothness and different persistence across the system

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Following an FFR innovation:

\vspace{0.15em}

\footnotesize
\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
IP falls under both estimators, with a larger long-horizon decline under LP

\vspace{0.15em}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
CPI initially rises under both, but the LP response subsequently declines toward zero while the VAR response continues rising

\end{itemize}

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} The price-puzzle pattern appears initially, but its persistence depends on the estimator


# Structural IRFs Using Local Projections


## Structural LP: Identification and Estimation {.t}

\vspace{0.5em}

\normalsize

\textcolor{navyblue}{$\bullet$} LPs accommodate recursive, long-run, sign and external-instrument identification

\vspace{0.35em}

\normalsize

\textcolor{navyblue}{$\bullet$} We adopt recursive identification with FFR last, matching the SVAR

\vspace{-0.6em}

\footnotesize
$$
\begin{aligned}
y_{j,t+h}={}&\alpha_{j,h}+\beta_{j,h}FFR_t
+\gamma_{j,h}'(IP_t,CPI_t)'\\
&+\sum_{\ell=1}^{14}\delta_{j,h,\ell}'y_{t-\ell}
+v_{j,t+h}^{(h)}
\end{aligned}
$$

\vspace{0.25em}

\normalsize

\textcolor{navyblue}{$\bullet$} Use \texttt{lp\_lin()} with Cholesky identification and horizon-specific regressions

\vspace{-0.5em}

\footnotesize
$$
\widehat{\Theta}_{j,3,h}^{LP}
=
\widehat{b}_{33}^{0}\widehat{\beta}_{j,h}
$$

\vspace{0.35em}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} Same identifying assumptions and shock normalization, but direct LP estimation

---

```{r}
#| include: false

# Recursive structural LP: IP, CPI, FFR
lp_struct <- lpirfs::lp_lin(
  endog_data = as.data.frame(x[, variables, drop = FALSE]),
  lags_endog_lin = p_aic + 1,
  trend = 0,
  shock_type = 1,
  confint = 1.96,
  hor = irf_horizon,
  num_cores = 1
)

# Extract responses to a unit FFR impact
mp_lp_unit <- t(
  lp_struct$irf_lin_mean[, , 3]
)

colnames(mp_lp_unit) <- variables

# Match recursive SVAR shock magnitude
mp_impact <- policy_svar_levels$FFR[1]

mp_lp_struct <- mp_lp_unit * mp_impact

# Responses in levels
policy_lp_struct_levels <- to_levels(
  mp_lp_struct
)

# Verify dimensions and correspondence with direct OLS
stopifnot(
  all(dim(mp_lp_unit) == c(irf_horizon + 1, 3))
)

lp_difference <- max(abs(
  mp_lp_unit - lp_irfs[, , "FFR"]
))

stopifnot(lp_difference < 1e-6)

```

## Structural IRFs: SVAR versus LP

```{r}
#| echo: false
#| fig-width: 9.0
#| fig-height: 3.7

plot_comparison(
  list(
    policy_svar_levels,
    policy_lp_struct_levels
  ),
  labels = c(
    "Recursive SVAR",
    "Structural LP"
  ),
  line_types = c(1, 2)
)
```

---

## Structural LP: Interpretation {.t}

\vspace{0.65em}

\normalsize

\textcolor{navyblue}{$\bullet$} Both estimators impose the same FFR-last recursive restrictions and agree on impact

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} LP responses exhibit different medium- and long-horizon dynamics relative to the SVAR

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} CPI initially rises under both methods, but the LP response subsequently weakens toward zero

\vspace{0.55em}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} Recursive identification does not remove the short-run price-puzzle pattern; its persistence depends on the estimator
