---
title: "Case Study 2 Part 1"
format: beamer
editor: visual
execute: 
  echo: true
  warning: false
  message: false
---

## Data Overview

\normalsize

\textcolor{blue}{$\bullet$} Data extracted from the Federal Reserve Economic Data (FRED)

\textcolor{blue}{$\bullet$} Selected series:

\footnotesize
\begin{itemize}
  \item[\raise0.15ex\hbox{\tiny$\circ$}]
    \textbf{INDPRO:} Industrial Production (Total Index)

  \item[\raise0.15ex\hbox{\tiny$\circ$}]
    \textbf{CPIAUCSL:} Consumer Price Index (All Urban Consumers)

  \item[\raise0.15ex\hbox{\tiny$\circ$}]
    \textbf{FEDFUNDS:} Effective Federal Funds Rate
\end{itemize}

\normalsize

\textcolor{blue}{$\bullet$} Original sample: January 1959 -- December 2019

\footnotesize
\begin{itemize}
  \item[\textcolor{blue}{$\rightarrow$}]
  \textbf{Effective sample: March 1959 -- December 2019}
  \quad \textit{(...Why?)}
\end{itemize}

\scriptsize

```{r}
#| echo: false
#| output: false
#| 
# Load the FRED-MD CSV.
raw <- read.csv("2026-rev-08-MD.csv", check.names = FALSE)

# The first row contains the FRED-MD transformation codes.
fred_codes <- unlist(
  raw[1, c("INDPRO", "CPIAUCSL", "FEDFUNDS")]
)

# Keep only the variables needed for the case.
dat <- raw[, c(
  "sasdate",
  "INDPRO",
  "CPIAUCSL",
  "FEDFUNDS"
)]

names(dat)[1] <- "Date"
```

```{r}
#| echo: false
#| output: false
# Convert dates.
dat$Date <- as.Date(
  dat$Date,
  format = "%m/%d/%Y"
)

# End the sample in December 2019.
dat <- dat[
  !is.na(dat$Date) &
    dat$Date <= as.Date("2019-12-31"),
]

dat <- dat[order(dat$Date), ]
```

---

## Economic Interpretation

\normalsize

\textcolor{blue}{$\bullet$} The three series capture distinct dimensions of aggregate economic activity

\textcolor{blue}{$\bullet$} Following \textit{Bernanke (2005)}, they can be interpreted as:

\footnotesize
\begin{itemize}
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
    \textit{INDPRO} -- Industrial Production
    $\rightarrow$ \textbf{Real activity / Output}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
    \textit{CPIAUCSL} -- Consumer Price Index
    $\rightarrow$ \textbf{Prices / Inflation}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
    \textit{FEDFUNDS} -- Effective Federal Funds Rate
    $\rightarrow$ \textbf{Monetary policy}
\end{itemize}

---

## Benchmark Transformations

\normalsize

\textcolor{blue}{$\bullet$} \textit{McCracken and Ng (2016)} provide benchmark transformations intended to obtain stationary representations of the FRED-MD series:

\footnotesize

\begin{center}
\renewcommand{\arraystretch}{1.4}

\begin{tabular}{lll}
Series & Transformation & Interpretation \\
\hline
\textit{INDPRO}
& $\Delta \log(\textit{INDPRO}_t)$
& Output growth \\

\textit{CPIAUCSL}
& $\Delta^2 \log(\textit{CPIAUCSL}_t)$
& Change in inflation \\

\textit{FEDFUNDS}
& $\Delta \textit{FEDFUNDS}_t$
& Change in the policy rate \\
\end{tabular}

\end{center}

\vspace{1em}

\normalsize

\textcolor{blue}{$\bullet$} These transformations are treated as benchmarks rather than assumptions

---

## Series in Levels {.t}

\vspace{1em}

\normalsize

\textcolor{blue}{$\bullet$} Before transforming the data, we inspect the original series over time

```{r}
#| output: false
#| echo: false

# FRED-MD transformations
dat_transformed <- data.frame(
  Date = dat$Date,

  INDPRO = c(
    NA,
    diff(log(dat$INDPRO))
  ),

  CPIAUCSL = c(
    NA, NA,
    diff(
      log(dat$CPIAUCSL),
      differences = 2
    )
  ),

  FEDFUNDS = c(
    NA,
    diff(dat$FEDFUNDS)
  )
)

# CPI transformation loses two observations.
# Use a common sample beginning in March 1959.
dat_common <- dat[-c(1, 2), ]
dat_transformed <- dat_transformed[-c(1, 2), ]

common_start <- c(
  as.integer(format(dat_common$Date[1], "%Y")),
  as.integer(format(dat_common$Date[1], "%m"))
)

# Series in levels
levels_ts <- ts(
  dat_common[, c(
    "INDPRO",
    "CPIAUCSL",
    "FEDFUNDS"
  )],
  start = common_start,
  frequency = 12
)

# Series after logs but before differencing
pre_diff_ts <- ts(
  cbind(
    log_INDPRO = log(dat_common$INDPRO),
    log_CPIAUCSL = log(dat_common$CPIAUCSL),
    FEDFUNDS = dat_common$FEDFUNDS
  ),
  start = common_start,
  frequency = 12
)

# FRED-MD transformed series
transformed_ts <- ts(
  dat_transformed[, c(
    "INDPRO",
    "CPIAUCSL",
    "FEDFUNDS"
  )],
  start = common_start,
  frequency = 12
)
```

---

## Series in Levels {.t}

```{r}
#| echo: false
#| fig-show: hold

plot(
  levels_ts[, "INDPRO"],
  main = "Industrial Production",
  xlab = "Year",
  ylab = "Index"
)

```

---

## Series in Levels {.t}

```{r}
#| echo: false
#| fig-show: hold

plot(
  levels_ts[, "CPIAUCSL"],
  main = "Consumer Price Index",
  xlab = "Year",
  ylab = "Index"
)

```

---

## Series in Levels {.t}

```{r}
#| echo: false
#| fig-show: hold

plot(
  levels_ts[, "FEDFUNDS"],
  main = "Federal Funds Rate",
  xlab = "Year",
  ylab = "Percent"
)

```

---

## Series in Levels {.t}

\vspace{1em}

\normalsize

\textcolor{blue}{$\bullet$} Before transforming the data, we inspect the original series over time

\small

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textit{INDPRO} exhibits a pronounced long-run upward trend and cyclical fluctuations

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textit{CPIAUCSL} displays a strong and persistent upward trend

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textit{FEDFUNDS} shows persistent movements and substantial variation

\end{itemize}

\vspace{1em}

\normalsize

\textcolor{blue}{$\rightarrow$} The visual evidence suggests that stationarity cannot be assumed in levels

---

## Transformed Series {.t}

\vspace{1em}

\normalsize

\textcolor{blue}{$\bullet$} Now, we apply the FRED-MD benchmark transformations:

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  $\Delta \log(\textit{INDPRO}_t)$

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  $\Delta^2 \log(\textit{CPIAUCSL}_t)$

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  $\Delta \textit{FEDFUNDS}_t$

\end{itemize}

---

## Transformed Series

```{r}
#| echo: false
#| fig-show: hold


plot(
  transformed_ts[, "INDPRO"],
  main = "Output Growth",
  xlab = "Year",
  ylab = expression(Delta~log(INDPRO))
)

```

---

## Transformed Series

```{r}
#| echo: false
#| fig-show: hold

plot(
  transformed_ts[, "CPIAUCSL"],
  main = "Change in Inflation",
  xlab = "Year",
  ylab = expression(Delta^2~log(CPI))
)

```

---

## Transformed Series

```{r}
#| echo: false
#| fig-show: hold

plot(
  transformed_ts[, "FEDFUNDS"],
  main = "Change in Policy Rate",
  xlab = "Year",
  ylab = expression(Delta~FEDFUNDS)
)
```

---

## Transformed Series {.t}

\vspace{1em}

\normalsize

\textcolor{blue}{$\bullet$} Now, we apply the FRED-MD benchmark transformations:

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  $\Delta \log(\textit{INDPRO}_t)$

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  $\Delta^2 \log(\textit{CPIAUCSL}_t)$

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  $\Delta \textit{FEDFUNDS}_t$

\end{itemize}

\normalsize

\textcolor{blue}{$\bullet$} The transformations substantially changes the persistence of the series \vspace{-0.5em} \small

\begin{itemize}
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
    Less persistence and no obvious deterministic trend
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
    Consistent with stationarity, but visual inspection is not sufficient to establish it
\end{itemize}

\normalsize

\textcolor{blue}{$\Rightarrow$} Still need to formally assess whether the transformed series are stationary

---

```{r}
#| echo: false


library(vars)

alpha <- 0.05

Y <- data.frame(
  Date = dat_transformed$Date,
  IP   = dat_transformed$INDPRO,
  CPI  = dat_transformed$CPIAUCSL,
  FFR  = dat_transformed$FEDFUNDS
)

x <- as.matrix(
  Y[, c("IP", "CPI", "FFR")]
)

stopifnot(all(is.finite(x)))


# Lag-order selection

# Allow up to two years of monthly lags
lag_max <- 24

lag_selection <- VARselect(
  x,
  lag.max = lag_max,
  type = "const"
)


# Store selected lag under each information criterion
lag_table <- data.frame(
  Criterion = c(
    "AIC",
    "HQ",
    "SC / BIC",
    "FPE"
  ),
  Selected_lag = as.integer(c(
    lag_selection$selection[["AIC(n)"]],
    lag_selection$selection[["HQ(n)"]],
    lag_selection$selection[["SC(n)"]],
    lag_selection$selection[["FPE(n)"]]
  ))
)


# Keep full criterion values
lag_criteria <- lag_selection$criteria


# Extract AIC- and BIC-selected lag orders

p_aic <- as.integer(
  lag_selection$selection[["AIC(n)"]]
)

p_bic <- as.integer(
  lag_selection$selection[["SC(n)"]]
)


# Estimate AIC- and BIC-selected VARs

fit_aic <- VAR(
  x,
  p = p_aic,
  type = "const"
)

fit_bic <- VAR(
  x,
  p = p_bic,
  type = "const"
)


# Residual serial-correlation horizon

# The Portmanteau test horizon: 2 years

serial_horizon <- 24


# Diagnostic function

run_var_diagnostics <- function(
  fit,
  serial_horizon
) {

  # VAR dynamic stability
  root_moduli <- roots(
    fit,
    modulus = TRUE
  )

  # Residual serial correlation
  serial_result <- serial.test(
    fit,
    lags.pt = serial_horizon,
    type = "PT.adjusted"
  )

  # Conditional heteroskedasticity
  arch_result <- arch.test(
    fit
  )

  # Multivariate residual normality
  normality_result <- normality.test(
    fit
  )

  # Collect results
  list(
    lag = fit$p,

    roots = root_moduli,

    stable = all(root_moduli < 1),

    max_root = max(root_moduli),

    serial_test = serial_result,

    serial_p = as.numeric(
      serial_result$serial$p.value
    ),

    arch_test = arch_result,

    arch_p = as.numeric(
      arch_result$arch.mul$p.value
    ),

    normality_test = normality_result,

    normality_p = as.numeric(
      normality_result$jb.mul$JB$p.value
    )
  )
}


# Run diagnostics

diag_aic <- run_var_diagnostics(
  fit_aic,
  serial_horizon
)

diag_bic <- run_var_diagnostics(
  fit_bic,
  serial_horizon
)


# Diagnostic comparison table

diagnostic_table <- data.frame(
  Criterion = c(
    "AIC",
    "SC / BIC"
  ),

  Selected_lag = c(
    diag_aic$lag,
    diag_bic$lag
  ),

  Stable = c(
    diag_aic$stable,
    diag_bic$stable
  ),

  Max_root = c(
    diag_aic$max_root,
    diag_bic$max_root
  ),

  Serial_p = c(
    diag_aic$serial_p,
    diag_bic$serial_p
  ),

  ARCH_p = c(
    diag_aic$arch_p,
    diag_bic$arch_p
  ),

  Normality_p = c(
    diag_aic$normality_p,
    diag_bic$normality_p
  )
)


# Add diagnostic decisions for presentation

diagnostic_table$Serial_correlation <- ifelse(
  diagnostic_table$Serial_p < alpha,
  "Rejected",
  "Not rejected"
)

diagnostic_table$ARCH_effects <- ifelse(
  diagnostic_table$ARCH_p < alpha,
  "Rejected",
  "Not rejected"
)

diagnostic_table$Normality <- ifelse(
  diagnostic_table$Normality_p < alpha,
  "Rejected",
  "Not rejected"
)

# Formatting helpers for slides

fmt_p <- function(p) {
  if (p < 0.0001) {
    "< 0.0001"
  } else {
    sprintf("%.4f", p)
  }
}

test_decision <- function(p, alpha = 0.05) {
  if (p < alpha) {
    "Reject"
  } else {
    "Do not reject"
  }
}

```

## VAR Specification

\normalsize

\textcolor{blue}{$\bullet$} Define the vector of stationary variables as

$$
\mathbf{y}_t =
\begin{bmatrix}
\Delta \log(\mathit{INDPRO}_t) \\
\Delta^2 \log(\mathit{CPIAUCSL}_t) \\
\Delta \mathit{FEDFUNDS}_t
\end{bmatrix}
=
\begin{bmatrix}
IP_t \\
CPI_t \\
FFR_t
\end{bmatrix}
$$

\textcolor{blue}{$\bullet$} Following \textit{Pfaff (2008)}, specify the reduced-form VAR($p$) as

$$
\mathbf{y}_t
=
\mathbf{c}
+
A_1 \mathbf{y}_{t-1}
+\cdots+
A_p \mathbf{y}_{t-p}
+
\mathbf{u}_t
$$

## Lag Order Selection

\normalsize

\textcolor{blue}{$\bullet$} Common methods to determine the lag order given by \textit{Kilian and Lütkepohl (2017)} include:

\vspace{0.4em}

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textbf{Top-down testing}: Start with many lags and remove the highest lag step by step if it adds little information
  \begin{itemize}
  \scriptsize
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]\textit{Adv.:} Simple and directly tests lag relevance
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]\textit{Lim.:} Depends on the starting order
  \end{itemize}
  \vspace{0.4em}

  \footnotesize
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textbf{Bottom-up testing}: Start with few lags and add more only if they meaningfully improve the model
  
  \begin{itemize}
  \scriptsize
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]\textit{Adv.:} Favors parsimonious models
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]\textit{Lim.:} May stop too early and miss longer-run dynamics
  \end{itemize}
  \vspace{0.4em}

  \footnotesize
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textbf{Forecast-based selection}: Choose the lag order with the best predictive performance
  \begin{itemize}
  \scriptsize
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]\textit{Adv.:} Directly targets forecasting \quad
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]\textit{Lim.:} Depends on horizon and evaluation sample
  \end{itemize}

\end{itemize}

---

## Information Criteria

\normalsize

\textcolor{blue}{$\bullet$} An alternative is to compare candidate VAR specifications using information criteria

\vspace{0.4em}

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textbf{AIC} -- Comparatively weaker penalty for additional parameters

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textbf{HQ} -- Intermediate penalty for model complexity

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textbf{SC/BIC} -- Stronger penalty, generally favoring more parsimonious models

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textbf{FPE} -- Evaluates the estimated prediction error of competing specifications

\end{itemize}

\vspace{0.4em}

\footnotesize

\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]\textit{Adv.:} Systematic comparison of fit and model complexity

\item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]\textit{Lim.:} Different criteria may select different lag orders
\end{itemize}

---

## Lag Selection Results -- Information Criteria

\normalsize

\textcolor{blue}{$\bullet$} Lag orders are selected using \texttt{VARselect()} from the \texttt{vars} package, comparing AIC, HQ, SC/BIC and FPE:

```{r}
#| echo: false
#| results: asis

cat("\\footnotesize\n")
cat("\\begin{center}\n")
cat("\\renewcommand{\\arraystretch}{1.4}\n\n")

cat("\\begin{tabular}{lc}\n")
cat("Criterion & Selected lag \\\\\n")
cat("\\hline\n")

for (i in seq_len(nrow(lag_table))) {
  cat(
    lag_table$Criterion[i],
    " & ",
    lag_table$Selected_lag[i],
    " \\\\\n",
    sep = ""
  )
}

cat("\\end{tabular}\n")
cat("\\end{center}\n")

```

\vspace{0.5em}

\normalsize

\textcolor{blue}{$\bullet$} AIC selects $p=14$, while SC/BIC selects $p=3$

\quad\textcolor{blue}{$\Rightarrow$} These define the two candidate lag specifications

---

## VAR Estimation

\normalsize

\textcolor{blue}{$\bullet$} Each candidate specification is estimated separately using \texttt{VAR()} from the \texttt{vars} package

\vspace{0.5em}

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  AIC specification: VAR(14)

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  SC/BIC specification: VAR(3)

\end{itemize}

\normalsize

\textcolor{blue}{$\bullet$} For each specification,

$$
\widehat{\mathbf{y}}_t
=
\widehat{\mathbf{c}}
+
\widehat{A}_1\mathbf{y}_{t-1}
+\cdots+
\widehat{A}_p\mathbf{y}_{t-p}
$$

\footnotesize

\quad\textcolor{blue}{$\rightarrow$} Each matrix $\widehat{A}_j$ contains the estimated own- and cross-lag effects among output growth, changes in inflation, and changes in the policy rate.

---

## Coefficient Interpretation

\vspace{-1em}

\normalsize

\textcolor{blue}{$\bullet$} Each row of $\widehat{A}_j$ corresponds to one VAR equation, while each column measures the effect of a lagged system variable

\vspace{0.5em}

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  Diagonal coefficients capture own-lag persistence

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  Off-diagonal coefficients capture cross-variable predictive effects

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  Coefficients are interpreted conditional on the remaining variables and included lags

\end{itemize}

\normalsize

\quad\textcolor{blue}{$\rightarrow$} The coefficients describe reduced-form dynamic relationships, not structural causal effects

---

## VAR Stability

\normalsize

\textcolor{blue}{$\bullet$} Stability assessed using \texttt{roots()} from the \texttt{vars} package

\vspace{0.4em}

\textcolor{blue}{$\bullet$} A VAR is stable if all characteristic roots satisfy $|\lambda_i| < 1$

\footnotesize

```{r}
#| echo: false
#| results: asis

stability_table <- data.frame(
  Specification = c(
    "AIC",
    "SC / BIC"
  ),
  Max_root = c(
    diag_aic$max_root,
    diag_bic$max_root
  )
)

stability_table$Max_root <- sprintf(
  "%.3f",
  stability_table$Max_root
)

cat("\\begin{center}\n")
cat("\\renewcommand{\\arraystretch}{1.4}\n")

cat("\\begin{tabular}{lc}\n")
cat("Specification & Maximum root modulus \\\\\n")
cat("\\hline\n")

for (i in seq_len(nrow(stability_table))) {
  cat(
    stability_table$Specification[i], " & ",
    stability_table$Max_root[i],
    " \\\\\n",
    sep = ""
  )
}

cat("\\end{tabular}\n")
cat("\\end{center}\n")

```

\normalsize

\textcolor{blue}{$\bullet$} Both specifications satisfy the stability condition

\footnotesize

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  AIC: maximum modulus $= 0.939$ -- dynamics are relatively persistent

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  SC/BIC: maximum modulus $= 0.665$ -- shocks decay more rapidly through the system

\end{itemize}

---

## Residual Diagnostics

\small

\textcolor{blue}{$\bullet$} Residual properties are assessed using \texttt{serial.test()}, \texttt{arch.test()} and \texttt{normality.test()} from the \texttt{vars} package \textit{(Pfaff, 2008)}

\vspace{0.4em}

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textbf{Serial correlation:}
  $H_0$ -- no residual serial correlation

  \scriptsize
  \begin{itemize}

    \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]
    AIC VAR(14): $p=0.0060$
    $\rightarrow$ reject $H_0$

    \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]
    SC/BIC VAR(3): $p<0.0001$
    $\rightarrow$ reject $H_0$

  \end{itemize}

  \scriptsize
  \quad Residual dependence remains under both lag specifications

  \vspace{0.35em}

  \footnotesize
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  \textbf{ARCH effects:}
  $H_0$ -- no conditional heteroskedasticity

  \scriptsize
  \begin{itemize}

    \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]
    Both specifications: $p<0.0001$
    $\rightarrow$ reject $H_0$

  \end{itemize}

  \scriptsize
  \quad Residual variance is not constant over time

  \vspace{0.35em}

  \footnotesize
  \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\circ$}}]
  (\textbf{Normality:}
  $H_0$ -- multivariate normality)

  \scriptsize
  \begin{itemize}

    \item[\raise0.15ex\hbox{\textcolor{blue}{\tiny$\blacktriangleright$}}]
    Both specifications: $p<0.0001$
    $\rightarrow$ reject $H_0$

  \end{itemize}

  \scriptsize
  \quad Normality is not required for VAR estimation; its rejection motivates
  bootstrap-based inference for the subsequent IRFs

\end{itemize}

---

## Model Selection

\normalsize

\textcolor{blue}{$\bullet$} Neither specification fully eliminates residual serial correlation

\vspace{0.5em}

\textcolor{blue}{$\bullet$} The SC/BIC VAR(3) is more parsimonious, while the AIC VAR(14) allows for richer monthly dynamics

\vspace{0.5em}

\textcolor{blue}{$\bullet$} VAR(14) is retained as the model specification for the subsequent IRF analysis

\vspace{0.25em}

\scriptsize

\begin{itemize}
  \item[\textcolor{blue}{$\rightarrow$}]\textbf{Results should be interpreted cautiously given the remaining residual misspecification!}
\end{itemize}

---
