---
title: "Case 2: Structural Vector AutoRegressive Models, Local Projections and Macro-economic Policy Analysis"
subtitle: "EBC4008 -- Time Series Analytics"
author:
  - "Can Idil"
  - "Sem Huizinga"
  - "Juan Diego Gaviño"
format:
  beamer:
    slide-level: 2
    header-includes:
      - |
        \definecolor{navyblue}{RGB}{10,32,72}

        \setbeamertemplate{navigation symbols}{}

        % Main presentation colors
        \setbeamercolor{normal text}{fg=navyblue,bg=white}
        \setbeamercolor{frametitle}{fg=navyblue}
        \setbeamercolor{structure}{fg=navyblue}

        % Header/footer colors
        \setbeamercolor{headlinebar}{bg=navyblue,fg=white}
        \setbeamercolor{footlinebar}{bg=navyblue,fg=white}

        % Header
        \setbeamertemplate{headline}{
          \leavevmode
          \begin{beamercolorbox}[
            wd=\paperwidth,
            ht=3.5ex,
            dp=1.0ex
          ]{headlinebar}

            \hbox to \paperwidth{%

              % Left: current section
              \rlap{%
                \hspace{1em}
                \scriptsize \insertsectionhead
              }%

              \hfill

              % Right
              \llap{%
                \scriptsize Case 2
                \hspace{1em}
              }%
            }

          \end{beamercolorbox}
        }

        % Footer
        \setbeamertemplate{footline}{
          \leavevmode
          \begin{beamercolorbox}[
            wd=\paperwidth,
            ht=2.8ex,
            dp=1.0ex
          ]{footlinebar}

            \hbox to \paperwidth{%

              % Left
              \rlap{%
                \hspace{1em}
                \tiny Maastricht University
              }%

              \hfill

              % Exact center
              \tiny EBC4008 -- Time Series Analytics

              \hfill

              % Right
              \llap{%
                \tiny \insertframenumber{} / \inserttotalframenumber
                \hspace{1em}
              }%
            }

          \end{beamercolorbox}
        }
editor: visual
execute:
  echo: true
  warning: false
  message: false
---

## Table of Contents {.t}

\normalsize

\begin{enumerate}
  \item Data Exploration
  \item Unit Root Testing
  \item VAR Specification and Model Selection
  \item Granger Causality
  \item Impulse Response Functions
\end{enumerate}


# Data Exploration
## Overview

\normalsize

\textcolor{navyblue}{$\bullet$} Data extracted from the Federal Reserve Economic Data (FRED)

\textcolor{navyblue}{$\bullet$} Selected series:

\footnotesize
\begin{itemize}
  \item[\raise0.15ex\hbox{\tiny$\circ$}]
    \textbf{INDPRO:} Industrial Production (Total Index)

  \item[\raise0.15ex\hbox{\tiny$\circ$}]
    \textbf{CPIAUCSL:} Consumer Price Index (All Urban Consumers)

  \item[\raise0.15ex\hbox{\tiny$\circ$}]
    \textbf{FEDFUNDS:} Effective Federal Funds Rate
\end{itemize}

\normalsize

\textcolor{navyblue}{$\bullet$} Original sample: January 1959 -- December 2019

\footnotesize
\begin{itemize}
  \item[\textcolor{navyblue}{$\rightarrow$}]
  \textbf{Effective sample: March 1959 -- December 2019}
  \quad \textit{(...Why?)}
\end{itemize}

\scriptsize

```{r}
#| echo: false
#| output: false
#| 
# Load the FRED-MD CSV.
raw <- read.csv("2026-rev-08-MD.csv", check.names = FALSE)

# The first row contains the FRED-MD transformation codes.
fred_codes <- unlist(
  raw[1, c("INDPRO", "CPIAUCSL", "FEDFUNDS")]
)

# Keep only the variables needed for the case.
dat <- raw[, c(
  "sasdate",
  "INDPRO",
  "CPIAUCSL",
  "FEDFUNDS"
)]

names(dat)[1] <- "Date"
```

```{r}
#| echo: false
#| output: false
# Convert dates.
dat$Date <- as.Date(
  dat$Date,
  format = "%m/%d/%Y"
)

# End the sample in December 2019.
dat <- dat[
  !is.na(dat$Date) &
    dat$Date <= as.Date("2019-12-31"),
]

dat <- dat[order(dat$Date), ]
```

------------------------------------------------------------------------

## Economic Interpretation

\normalsize

\textcolor{navyblue}{$\bullet$} The three series aim to capture distinct dimensions of aggregate economic activity

\textcolor{navyblue}{$\bullet$} Following \textit{Bernanke (2005)}, they can be interpreted as:

\footnotesize
\begin{itemize}
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
    \textit{INDPRO} -- Industrial Production
    $\rightarrow$ \textbf{Real activity / Output}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
    \textit{CPIAUCSL} -- Consumer Price Index
    $\rightarrow$ \textbf{Prices / Inflation}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
    \textit{FEDFUNDS} -- Effective Federal Funds Rate
    $\rightarrow$ \textbf{Monetary policy}
\end{itemize}

------------------------------------------------------------------------

## Benchmark Transformations

\normalsize

\textcolor{navyblue}{$\bullet$} \textit{McCracken and Ng (2016)} provide benchmark transformations intended to obtain stationary representations of the FRED-MD series:

\footnotesize

\begin{center}
\renewcommand{\arraystretch}{1.4}

\begin{tabular}{lll}
Series & Transformation & Interpretation \\
\hline
\textit{INDPRO}
& $\Delta \log(\textit{INDPRO}_t)$
& Output growth \\

\textit{CPIAUCSL}
& $\Delta^2 \log(\textit{CPIAUCSL}_t)$
& Change in inflation \\

\textit{FEDFUNDS}
& $\Delta \textit{FEDFUNDS}_t$
& Change in the policy rate \\
\end{tabular}

\end{center}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} These transformations are treated as benchmarks rather than assumptions

------------------------------------------------------------------------

## Series in Levels {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Before transforming the data, we inspect the original series over time

```{r}
#| output: false
#| echo: false

# FRED-MD transformations
dat_transformed <- data.frame(
  Date = dat$Date,

  INDPRO = c(
    NA,
    diff(log(dat$INDPRO))
  ),

  CPIAUCSL = c(
    NA, NA,
    diff(
      log(dat$CPIAUCSL),
      differences = 2
    )
  ),

  FEDFUNDS = c(
    NA,
    diff(dat$FEDFUNDS)
  )
)

# CPI transformation loses two observations.
# Use a common sample beginning in March 1959.
dat_common <- dat[-c(1, 2), ]
dat_transformed <- dat_transformed[-c(1, 2), ]

common_start <- c(
  as.integer(format(dat_common$Date[1], "%Y")),
  as.integer(format(dat_common$Date[1], "%m"))
)

# Series in levels
levels_ts <- ts(
  dat_common[, c(
    "INDPRO",
    "CPIAUCSL",
    "FEDFUNDS"
  )],
  start = common_start,
  frequency = 12
)

# Series after logs but before differencing
pre_diff_ts <- ts(
  cbind(
    log_INDPRO = log(dat_common$INDPRO),
    log_CPIAUCSL = log(dat_common$CPIAUCSL),
    FEDFUNDS = dat_common$FEDFUNDS
  ),
  start = common_start,
  frequency = 12
)

# FRED-MD transformed series
transformed_ts <- ts(
  dat_transformed[, c(
    "INDPRO",
    "CPIAUCSL",
    "FEDFUNDS"
  )],
  start = common_start,
  frequency = 12
)
```

------------------------------------------------------------------------

## Series in Levels {.t}

```{r}
#| echo: false
#| fig-show: hold

plot(
  levels_ts[, "INDPRO"],
  main = "Industrial Production",
  xlab = "Year",
  ylab = "Index"
)

```

------------------------------------------------------------------------

## Series in Levels {.t}

```{r}
#| echo: false
#| fig-show: hold

plot(
  levels_ts[, "CPIAUCSL"],
  main = "Consumer Price Index",
  xlab = "Year",
  ylab = "Index"
)

```

------------------------------------------------------------------------

## Series in Levels {.t}

```{r}
#| echo: false
#| fig-show: hold

plot(
  levels_ts[, "FEDFUNDS"],
  main = "Federal Funds Rate",
  xlab = "Year",
  ylab = "Percent"
)

```

------------------------------------------------------------------------

## Series in Levels {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Before transforming the data, we inspect the original series over time

\small

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textit{INDPRO} exhibits a pronounced long-run upward trend and cyclical fluctuations

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textit{CPIAUCSL} displays a strong and persistent upward trend

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textit{FEDFUNDS} shows substantial variation across the sample period

\end{itemize}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\rightarrow$} The visual evidence suggests that stationarity cannot be assumed in levels

------------------------------------------------------------------------

## Transformed Series {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Now, we apply the FRED-MD benchmark transformations:

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  $\Delta \log(\textit{INDPRO}_t)$

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  $\Delta^2 \log(\textit{CPIAUCSL}_t)$

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  $\Delta \textit{FEDFUNDS}_t$

\end{itemize}

------------------------------------------------------------------------

## Transformed Series

```{r}
#| echo: false
#| fig-show: hold


plot(
  transformed_ts[, "INDPRO"],
  main = "Output Growth",
  xlab = "Year",
  ylab = expression(Delta~log(INDPRO))
)

```

------------------------------------------------------------------------

## Transformed Series

```{r}
#| echo: false
#| fig-show: hold

plot(
  transformed_ts[, "CPIAUCSL"],
  main = "Change in Inflation",
  xlab = "Year",
  ylab = expression(Delta^2~log(CPI))
)

```

------------------------------------------------------------------------

## Transformed Series

```{r}
#| echo: false
#| fig-show: hold

plot(
  transformed_ts[, "FEDFUNDS"],
  main = "Change in Policy Rate",
  xlab = "Year",
  ylab = expression(Delta~FEDFUNDS)
)
```

------------------------------------------------------------------------

## Transformed Series {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Now, we apply the FRED-MD benchmark transformations:

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  $\Delta \log(\textit{INDPRO}_t)$

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  $\Delta^2 \log(\textit{CPIAUCSL}_t)$

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  $\Delta \textit{FEDFUNDS}_t$

\end{itemize}

\normalsize

\textcolor{navyblue}{$\bullet$} The transformations substantially changes the persistence of the series \vspace{-0.5em} \small

\begin{itemize}
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
    Less persistence and no obvious deterministic trend
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
    Consistent with stationarity, but visual inspection is not sufficient to establish it
\end{itemize}

\normalsize

\textcolor{navyblue}{$\Rightarrow$} Still need to formally assess whether the transformed series are stationary


```{r}
#| include: false

# Unit root testing

library(bootUR)

IP  <- log(dat$INDPRO)
CPI <- log(dat$CPIAUCSL)
FFR <- dat$FEDFUNDS

set.seed(123)
B <- 4999
alpha <- 0.05


# Automatic order of integration

IP_trend <- order_integration(
  IP,
  method = "boot_adf",
  deterministics = "trend",
  B = B
)

CPI_trend <- order_integration(
  CPI,
  method = "boot_adf",
  deterministics = "trend",
  B = B
)

FFR_trend <- order_integration(
  FFR,
  method = "boot_adf",
  deterministics = "trend",
  B = B
)


IP_intercept <- order_integration(
  IP,
  method = "boot_adf",
  deterministics = "intercept",
  B = B
)

CPI_intercept <- order_integration(
  CPI,
  method = "boot_adf",
  deterministics = "intercept",
  B = B
)

FFR_intercept <- order_integration(
  FFR,
  method = "boot_adf",
  deterministics = "intercept",
  B = B
)


# Compare deterministic specifications

unit_root_table <- data.frame(
  Series = c("IP", "CPI", "FFR"),

  Trend = c(
    IP_trend$order_int,
    CPI_trend$order_int,
    FFR_trend$order_int
  ),

  Intercept = c(
    IP_intercept$order_int,
    CPI_intercept$order_int,
    FFR_intercept$order_int
  )
)



# Manual bootstrap ADF tests

# IP
ur_IP_0 <- boot_adf(
  IP,
  deterministics = "trend",
  B = B
)

ur_IP_1 <- boot_adf(
  diff(IP),
  deterministics = "intercept",
  B = B
)

ur_IP_2 <- boot_adf(
  diff(diff(IP)),
  deterministics = "intercept",
  B = B
)


# CPI
ur_CPI_0 <- boot_adf(
  CPI,
  deterministics = "trend",
  B = B
)

ur_CPI_1 <- boot_adf(
  diff(CPI),
  deterministics = "intercept",
  B = B
)

ur_CPI_2 <- boot_adf(
  diff(diff(CPI)),
  deterministics = "intercept",
  B = B
)


# FFR
ur_FFR_0 <- boot_adf(
  FFR,
  deterministics = "trend",
  B = B
)

ur_FFR_1 <- boot_adf(
  diff(FFR),
  deterministics = "intercept",
  B = B
)

ur_FFR_2 <- boot_adf(
  diff(diff(FFR)),
  deterministics = "intercept",
  B = B
)


# Manual results

unit_root_table_manual <- data.frame(
  Series = c("IP", "CPI", "FFR"),

  Level = c(
    ur_IP_0$p.value,
    ur_CPI_0$p.value,
    ur_FFR_0$p.value
  ),

  Delta_1 = c(
    ur_IP_1$p.value,
    ur_CPI_1$p.value,
    ur_FFR_1$p.value
  ),

  Delta_2 = c(
    ur_IP_2$p.value,
    ur_CPI_2$p.value,
    ur_FFR_2$p.value
  )
)


# Infer order of integration from manual tests

unit_root_table_manual$Order <- ifelse(
  unit_root_table_manual$Level < alpha,
  0,
  ifelse(
    unit_root_table_manual$Delta_1 < alpha,
    1,
    2
  )
)



# Tables for slides

unit_root_order_slide <- unit_root_table

unit_root_manual_slide <- unit_root_table_manual

unit_root_manual_slide$Level <- format.pval(
  unit_root_manual_slide$Level,
  digits = 3,
  eps = 0.001
)

unit_root_manual_slide$Delta_1 <- format.pval(
  unit_root_manual_slide$Delta_1,
  digits = 3,
  eps = 0.001
)

unit_root_manual_slide$Delta_2 <- format.pval(
  unit_root_manual_slide$Delta_2,
  digits = 3,
  eps = 0.001
)
```

# Unit Root Testing
## Integrated processes


\normalsize

\textcolor{navyblue}{$\bullet$} The objective is to determine the order of integration of each series $$y_t \sim I(d)$$ where $d$ is the number of differences required to obtain a stationary series

\vspace{0.45em}

\textcolor{navyblue}{$\bullet$} This determines the transformation used before estimating the VAR.

---

## Manual Verification {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} First, use \texttt{boot\_adf()} directly to the level and successive differences $$\Delta^2 y_t\rightarrow \Delta y_t\rightarrow y_t$$


\vspace{0.35em}

\scriptsize

```{r}
#| echo: false
#| results: asis

knitr::kable(
  unit_root_manual_slide,
  row.names = FALSE,
  align = "c"
)

```

\vspace{0.25em}

\textcolor{navyblue}{$\bullet$} $p<0.05$ rejects the unit-root null; the integration order is the number of differences required for rejection

------------------------------------------------------------------------

## Automatic Integration Order {.t}

\normalsize

\textcolor{navyblue}{$\bullet$} We then apply  \texttt{order\_integration()} with bootstrap ADF tests

\vspace{0.3em}

\footnotesize

\texttt{order\_integration(}

\texttt{\hspace{1em} x, method = "boot\_adf",}

\texttt{\hspace{1em} deterministics = "...", B = 4999)}

\vspace{0.35em}

\normalsize

\textcolor{navyblue}{$\bullet$} The procedure tests successive integration orders automatically

\vspace{0.3em}

\scriptsize

```{r}
#| echo: false
#| results: asis

knitr::kable(
  unit_root_order_slide,
  row.names = FALSE,
  align = "c"
)

```

\vspace{0.3em}
\footnotesize

\quad\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}} The procedure is repeated with a trend and with an intercept to assess sensitivity to the deterministic specification.

------------------------------------------------------------------------

## CPI: I(1) or I(2)?{.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} CPI is sensitive to the deterministic specification

\vspace{0.35em}

\footnotesize

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \texttt{order\_integration()} with trend: $I(2)$

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \texttt{order\_integration()} with intercept: $I(1)$

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  Manual \texttt{boot\_adf()} sequence: $I(1)$ at the 5\% level

\end{itemize}

\vspace{0.4em}

\normalsize

\textcolor{navyblue}{$\bullet$} The integration order of CPI is therefore not fully robust

\vspace{0.3em}

\textcolor{navyblue}{$\bullet$} We retain the FRED-MD benchmark transformation $\Delta^2\log(CPI)$ for the VAR


```{r}
#| echo: false


library(vars)

alpha <- 0.05

Y <- data.frame(
  Date = dat_transformed$Date,
  IP   = dat_transformed$INDPRO,
  CPI  = dat_transformed$CPIAUCSL,
  FFR  = dat_transformed$FEDFUNDS
)

x <- as.matrix(
  Y[, c("IP", "CPI", "FFR")]
)

stopifnot(all(is.finite(x)))


# Lag-order selection

# Allow up to two years of monthly lags
lag_max <- 24

lag_selection <- VARselect(
  x,
  lag.max = lag_max,
  type = "const"
)


# Store selected lag under each information criterion
lag_table <- data.frame(
  Criterion = c(
    "AIC",
    "HQ",
    "SC / BIC",
    "FPE"
  ),
  Selected_lag = as.integer(c(
    lag_selection$selection[["AIC(n)"]],
    lag_selection$selection[["HQ(n)"]],
    lag_selection$selection[["SC(n)"]],
    lag_selection$selection[["FPE(n)"]]
  ))
)


# Keep full criterion values
lag_criteria <- lag_selection$criteria


# Extract AIC- and BIC-selected lag orders

p_aic <- as.integer(
  lag_selection$selection[["AIC(n)"]]
)

p_bic <- as.integer(
  lag_selection$selection[["SC(n)"]]
)


# Estimate AIC- and BIC-selected VARs

fit_aic <- VAR(
  x,
  p = p_aic,
  type = "const"
)

fit_bic <- VAR(
  x,
  p = p_bic,
  type = "const"
)


# Residual serial-correlation horizon

# The Portmanteau test horizon: 2 years

serial_horizon <- 24


# Diagnostic function

run_var_diagnostics <- function(
  fit,
  serial_horizon
) {

  # VAR dynamic stability
  root_moduli <- roots(
    fit,
    modulus = TRUE
  )

  # Residual serial correlation
  serial_result <- serial.test(
    fit,
    lags.pt = serial_horizon,
    type = "PT.adjusted"
  )

  # Conditional heteroskedasticity
  arch_result <- arch.test(
    fit
  )

  # Multivariate residual normality
  normality_result <- normality.test(
    fit
  )

  # Collect results
  list(
    lag = fit$p,

    roots = root_moduli,

    stable = all(root_moduli < 1),

    max_root = max(root_moduli),

    serial_test = serial_result,

    serial_p = as.numeric(
      serial_result$serial$p.value
    ),

    arch_test = arch_result,

    arch_p = as.numeric(
      arch_result$arch.mul$p.value
    ),

    normality_test = normality_result,

    normality_p = as.numeric(
      normality_result$jb.mul$JB$p.value
    )
  )
}


# Run diagnostics

diag_aic <- run_var_diagnostics(
  fit_aic,
  serial_horizon
)

diag_bic <- run_var_diagnostics(
  fit_bic,
  serial_horizon
)


# Diagnostic comparison table

diagnostic_table <- data.frame(
  Criterion = c(
    "AIC",
    "SC / BIC"
  ),

  Selected_lag = c(
    diag_aic$lag,
    diag_bic$lag
  ),

  Stable = c(
    diag_aic$stable,
    diag_bic$stable
  ),

  Max_root = c(
    diag_aic$max_root,
    diag_bic$max_root
  ),

  Serial_p = c(
    diag_aic$serial_p,
    diag_bic$serial_p
  ),

  ARCH_p = c(
    diag_aic$arch_p,
    diag_bic$arch_p
  ),

  Normality_p = c(
    diag_aic$normality_p,
    diag_bic$normality_p
  )
)


# Add diagnostic decisions for presentation

diagnostic_table$Serial_correlation <- ifelse(
  diagnostic_table$Serial_p < alpha,
  "Rejected",
  "Not rejected"
)

diagnostic_table$ARCH_effects <- ifelse(
  diagnostic_table$ARCH_p < alpha,
  "Rejected",
  "Not rejected"
)

diagnostic_table$Normality <- ifelse(
  diagnostic_table$Normality_p < alpha,
  "Rejected",
  "Not rejected"
)

# Formatting helpers for slides

fmt_p <- function(p) {
  if (p < 0.0001) {
    "< 0.0001"
  } else {
    sprintf("%.4f", p)
  }
}

test_decision <- function(p, alpha = 0.05) {
  if (p < alpha) {
    "Reject"
  } else {
    "Do not reject"
  }
}

```

# VAR Specification and Model Selection
## VAR Specification {.t}

\normalsize

\textcolor{navyblue}{$\bullet$} Define the vector of stationary variables as

$$
\mathbf y_t
=
\begin{bmatrix}
IP_t\\
CPI_t\\
FFR_t
\end{bmatrix}
=
\begin{bmatrix}
\Delta\log(INDPRO_t)\\
\Delta^2\log(CPIAUCSL_t)\\
\Delta FEDFUNDS_t
\end{bmatrix}
$$

\vspace{0.3em}

\textcolor{navyblue}{$\bullet$} The VAR($p$) can be written in companion form as

\scriptsize

$$
\boldsymbol{\xi}_t
=
\mathbf c^*
+
\mathbf F \boldsymbol{\xi}_{t-1}
+
\mathbf v_t
$$

$$
\boldsymbol{\xi}_t
=
\begin{bmatrix}
\mathbf y_t\\
\mathbf y_{t-1}\\
\vdots\\
\mathbf y_{t-p+1}
\end{bmatrix},
\quad
\mathbf v_t
=
\begin{bmatrix}
\mathbf u_t\\
\mathbf 0\\
\vdots\\
\mathbf 0
\end{bmatrix},
\quad
\mathbf F
=
\begin{bmatrix}
A_1 & A_2 & \cdots & A_{p-1} & A_p\\
I_3 & 0   & \cdots & 0       & 0\\
0   & I_3 & \cdots & 0       & 0\\
\vdots & \vdots & \ddots & \vdots & \vdots\\
0   & 0   & \cdots & I_3     & 0
\end{bmatrix}
$$
\footnotesize

\vspace{-0.5em}

\qquad $\rightarrow$ Here, $\boldsymbol{\xi}_t$ stacks the current and lagged values of the system

---

## Lag Order Selection

\normalsize

\textcolor{navyblue}{$\bullet$} Common methods to determine the lag order given by \textit{Kilian and Lütkepohl (2017)} include:

\vspace{0.4em}

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textbf{Top-down testing}: Start with many lags and remove the highest lag step by step if it adds little information
  \begin{itemize}
  \scriptsize
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]\textit{Adv.:} Simple and directly tests lag relevance
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]\textit{Lim.:} Depends on the starting order
  \end{itemize}
  \vspace{0.4em}

  \footnotesize
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textbf{Bottom-up testing}: Start with few lags and add more only if they meaningfully improve the model
  
  \begin{itemize}
  \scriptsize
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]\textit{Adv.:} Favors parsimonious models
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]\textit{Lim.:} May stop too early and miss longer-run dynamics
  \end{itemize}
  \vspace{0.4em}

  \footnotesize
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textbf{Forecast-based selection}: Choose the lag order with the best predictive performance
  \begin{itemize}
  \scriptsize
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]\textit{Adv.:} Directly targets forecasting \quad
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]\textit{Lim.:} Depends on horizon and evaluation sample
  \end{itemize}

\end{itemize}

------------------------------------------------------------------------

## Information Criteria{.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} An alternative is to compare candidate VAR specifications using information criteria

\vspace{0.4em}

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textbf{AIC} -- Comparatively weaker penalty for additional parameters

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textbf{HQ} -- Intermediate penalty for model complexity

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textbf{SC/BIC} -- Stronger penalty, generally favoring more parsimonious models

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textbf{FPE} -- Evaluates the estimated prediction error of competing specifications

\end{itemize}

\vspace{0.4em}

\footnotesize

\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]\textit{Adv.:} Systematic comparison of fit and model complexity

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]\textit{Lim.:} Different criteria may select different lag orders
\end{itemize}

------------------------------------------------------------------------

## Lag Selection Results -- Information Criteria{.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Lag orders are selected using \texttt{VARselect()} from the \texttt{vars} package, comparing AIC, HQ, SC/BIC and FPE:

```{r}
#| echo: false
#| results: asis

cat("\\footnotesize\n")
cat("\\begin{center}\n")
cat("\\renewcommand{\\arraystretch}{1.4}\n\n")

cat("\\begin{tabular}{lc}\n")
cat("Criterion & Selected lag \\\\\n")
cat("\\hline\n")

for (i in seq_len(nrow(lag_table))) {
  cat(
    lag_table$Criterion[i],
    " & ",
    lag_table$Selected_lag[i],
    " \\\\\n",
    sep = ""
  )
}

cat("\\end{tabular}\n")
cat("\\end{center}\n")

```

\vspace{0.5em}

\normalsize

\textcolor{navyblue}{$\bullet$} AIC selects $p=14$, while SC/BIC selects $p=3$

\quad\textcolor{navyblue}{$\Rightarrow$} These define the two candidate lag specifications

------------------------------------------------------------------------

## VAR Estimation {.t}

\normalsize

\textcolor{navyblue}{$\bullet$} The two candidate specifications are estimated using \texttt{VAR()}

\vspace{0.25em}

\footnotesize

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  AIC specification: VAR(14)

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  SC/BIC specification: VAR(3)

\end{itemize}

\vspace{0.3em}

\normalsize

\textcolor{navyblue}{$\bullet$} For each lag $j$, estimation produces

$$
\widehat A_j =
\begin{bmatrix}
\hat a_{11,j} & \hat a_{12,j} & \hat a_{13,j}\\
\hat a_{21,j} & \hat a_{22,j} & \hat a_{23,j}\\
\hat a_{31,j} & \hat a_{32,j} & \hat a_{33,j}
\end{bmatrix}
$$

\vspace{0.35em}

\footnotesize

\quad \raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}} Rows correspond to the $IP$, $CPI$ and $FFR$ equations; columns correspond to lagged $IP$, $CPI$ and $FFR$

\vspace{0.3em}

\quad \raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}} Diagonal elements capture own-lag effects, while off-diagonal elements capture cross-variable predictive effects

---

## VAR Stability {.t}

\vspace{1em}

\small

\textcolor{navyblue}{$\bullet$} Stability is assessed using \texttt{roots()} from the \texttt{vars} package

\vspace{0.4em}

\textcolor{navyblue}{$\bullet$} A VAR is stable if all eigenvalues of the companion matrix satisfy \scriptsize$|\lambda_i|< 1$

\footnotesize

```{r}
#| echo: false
#| results: asis

stability_table <- data.frame(
  Specification = c(
    "AIC",
    "SC / BIC"
  ),
  Max_root = c(
    diag_aic$max_root,
    diag_bic$max_root
  )
)

stability_table$Max_root <- sprintf(
  "%.3f",
  stability_table$Max_root
)

cat("\\begin{center}\n")
cat("\\renewcommand{\\arraystretch}{1.4}\n")

cat("\\begin{tabular}{lc}\n")
cat("Specification & Maximum root modulus \\\\\n")
cat("\\hline\n")

for (i in seq_len(nrow(stability_table))) {
  cat(
    stability_table$Specification[i], " & ",
    stability_table$Max_root[i],
    " \\\\\n",
    sep = ""
  )
}

cat("\\end{tabular}\n")
cat("\\end{center}\n")

```

\small
\textcolor{navyblue}{$\bullet$} Both specifications satisfy the stability condition
\vspace{0.25em}
\footnotesize
\begin{itemize}
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  AIC: maximum modulus $= 0.939$ -- closer to the unit-circle boundary
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  SC/BIC: maximum modulus $= 0.665$ -- farther inside the unit circle
\end{itemize}
\vspace{0.25em}
\small
\textcolor{navyblue}{$\bullet$} Hence, both estimated VARs are dynamically stable, although the AIC specification lies closer to the stability boundary

---

## Residual Diagnostics

\small

\textcolor{navyblue}{$\bullet$} Residual properties are assessed using \texttt{serial.test()}, \texttt{arch.test()} and \texttt{normality.test()} from the \texttt{vars} package \textit{(Pfaff, 2008)}

\vspace{0.4em}

\footnotesize
\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textbf{Serial correlation:}
  $H_0$ -- no residual serial correlation

  \scriptsize
  \begin{itemize}

    \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]
    AIC VAR(14): $p=0.0060$
    $\rightarrow$ reject $H_0$

    \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]
    SC/BIC VAR(3): $p<0.0001$
    $\rightarrow$ reject $H_0$

  \end{itemize}

  \scriptsize
  \quad Residual dependence remains under both lag specifications

  \vspace{0.35em}

  \footnotesize
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  \textbf{ARCH effects:}
  $H_0$ -- no conditional heteroskedasticity

  \scriptsize
  \begin{itemize}

    \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]
    Both specifications: $p<0.0001$
    $\rightarrow$ reject $H_0$

  \end{itemize}

  \scriptsize
  \quad Residual variance is not constant over time

  \vspace{0.35em}

  \footnotesize
  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  (\textbf{Normality:}
  $H_0$ -- multivariate normality)

  \scriptsize
  \begin{itemize}

    \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\blacktriangleright$}}]
    Both specifications: $p<0.0001$
    $\rightarrow$ reject $H_0$

  \end{itemize}

  \scriptsize
  \quad Normality is not required for VAR estimation; its rejection motivates
  bootstrap-based inference for the subsequent IRFs

\end{itemize}

------------------------------------------------------------------------

## Model Selection

\normalsize

\textcolor{navyblue}{$\bullet$} Neither specification fully eliminates residual serial correlation

\vspace{0.5em}

\textcolor{navyblue}{$\bullet$} The SC/BIC VAR(3) is more parsimonious, while the AIC VAR(14) allows for richer monthly dynamics

\vspace{0.5em}

\textcolor{navyblue}{$\bullet$} VAR(14) is retained as the model specification for the subsequent IRF analysis

\vspace{0.25em}

\scriptsize

\begin{itemize}
  \item[\textcolor{navyblue}{$\rightarrow$}]\textbf{Results should be interpreted cautiously given the remaining residual misspecification!}
\end{itemize}


```{r}
#| include: false

#install.packages("car")
library(car)

# Granger causality

#alpha <- 0.05


# CPI and FFR -> IP
gc_IP <- causality(
  fit_aic,
  cause = c("CPI", "FFR")
)$Granger

# IP and FFR -> CPI
gc_CPI <- causality(
  fit_aic,
  cause = c("IP", "FFR")
)$Granger

# IP and CPI -> FFR
gc_FFR <- causality(
  fit_aic,
  cause = c("IP", "CPI")
)$Granger


# Collect the three block-test results
granger_block_table <- data.frame(
  x = c(
    "CPI + FFR",
    "IP + FFR",
    "IP + CPI"
  ),

  y = c(
    "IP",
    "CPI",
    "FFR"
  ),

  F_statistic = c(
    as.numeric(gc_IP$statistic),
    as.numeric(gc_CPI$statistic),
    as.numeric(gc_FFR$statistic)
  ),

  p_value = c(
    gc_IP$p.value,
    gc_CPI$p.value,
    gc_FFR$p.value
  )
)

granger_block_table$Decision <- ifelse(
  granger_block_table$p_value < alpha,
  "Reject H0",
  "Do not reject H0"
)



# Six (conditional) pairwise tests


# IP -> CPI
#
# H0:
# IP.l1 = IP.l2 = ... = IP.l14 = 0
# in the CPI equation


test_IP_CPI <- linearHypothesis(
  fit_aic$varresult[["CPI"]],
  paste0(
    "IP.l",
    1:p_aic,
    " = 0"
  )
)


# IP -> FFR

test_IP_FFR <- linearHypothesis(
  fit_aic$varresult[["FFR"]],
  paste0(
    "IP.l",
    1:p_aic,
    " = 0"
  )
)


# CPI -> IP

test_CPI_IP <- linearHypothesis(
  fit_aic$varresult[["IP"]],
  paste0(
    "CPI.l",
    1:p_aic,
    " = 0"
  )
)


# CPI -> FFR

test_CPI_FFR <- linearHypothesis(
  fit_aic$varresult[["FFR"]],
  paste0(
    "CPI.l",
    1:p_aic,
    " = 0"
  )
)



# FFR -> IP

test_FFR_IP <- linearHypothesis(
  fit_aic$varresult[["IP"]],
  paste0(
    "FFR.l",
    1:p_aic,
    " = 0"
  )
)



# FFR -> CPI

test_FFR_CPI <- linearHypothesis(
  fit_aic$varresult[["CPI"]],
  paste0(
    "FFR.l",
    1:p_aic,
    " = 0"
  )
)


# Collect results

granger_pairwise_table <- data.frame(
  x = c(
    "IP",
    "IP",
    "CPI",
    "CPI",
    "FFR",
    "FFR"
  ),

  y = c(
    "CPI",
    "FFR",
    "IP",
    "FFR",
    "IP",
    "CPI"
  ),

  F_statistic = c(
    test_IP_CPI$F[2],
    test_IP_FFR$F[2],
    test_CPI_IP$F[2],
    test_CPI_FFR$F[2],
    test_FFR_IP$F[2],
    test_FFR_CPI$F[2]
  ),

  p_value = c(
    test_IP_CPI$`Pr(>F)`[2],
    test_IP_FFR$`Pr(>F)`[2],
    test_CPI_IP$`Pr(>F)`[2],
    test_CPI_FFR$`Pr(>F)`[2],
    test_FFR_IP$`Pr(>F)`[2],
    test_FFR_CPI$`Pr(>F)`[2]
  )
)


# Decision at the 5% level
granger_pairwise_table$Decision <- ifelse(
  granger_pairwise_table$p_value < alpha,
  "Reject H0",
  "Do not reject H0"
)



# Tables

granger_block_slide <- granger_block_table

granger_block_slide$F_statistic <- round(
  granger_block_slide$F_statistic,
  2
)

granger_block_slide$p_value <- format.pval(
  granger_block_slide$p_value,
  digits = 3,
  eps = 0.001
)


granger_pairwise_slide <- granger_pairwise_table

granger_pairwise_slide$F_statistic <- round(
  granger_pairwise_slide$F_statistic,
  2
)

granger_pairwise_slide$p_value <- format.pval(
  granger_pairwise_slide$p_value,
  digits = 3,
  eps = 0.001
)

```

# Granger Causality
## Granger Causality {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} $x$ Granger-causes $y$ if past $x$ improves the prediction of $y$, conditional on the VAR

\vspace{0.45em}

\textcolor{navyblue}{$\bullet$} Granger causality is predictive, not necessarily structural causality

\vspace{0.45em}

\textcolor{navyblue}{$\bullet$} The \texttt{vars} package provides the function \texttt{causality()}

\vspace{0.4em}

\footnotesize
\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}] For a some variable (or vector in this case) $x$,

\end{itemize}

$$H_0:
\beta_{x_{1},1}
=
\cdots
=
\beta_{x_{1},p}
=
\beta_{x_{2},1}
=
\cdots
=\beta_{x_{2},p}
=
0$$

$$H_1:
\beta_{x_{i},j}\neq0
\quad
\text{for at least one }j$$

------------------------------------------------------------------------

## Granger Causality

\normalsize

\textcolor{navyblue}{$\bullet$} In a trivariate VAR, \texttt{causality()} tests the chosen variables against the remaining one jointly

\vspace{0.35em}

\small

```{r}
#| echo: false
#| results: asis

knitr::kable(
  granger_block_slide,
  row.names = FALSE,
  align = "c"
)
```

------------------------------------------------------------------------

## Pairwise Granger Causality {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Test the six directional relationships separately while retaining the full trivariate VAR

\vspace{0.3em}

\small

```{r}
#| echo: false
#| results: asis

knitr::kable(
  granger_pairwise_slide,
  row.names = FALSE,
  align = "c"
)

```

\vspace{-1em}
\scriptsize

$$H_0:
\beta_{x,1}
=
\cdots
=
\beta_{x,p}
=
0$$

$$H_1:
\beta_{x,j}\neq0
\quad
\text{for at least one }j$$


```{r}
#| include: false

# IRF settings

irf_horizon <- 48
irf_runs <- 4999
irf_ci <- 0.95

variables <- c(
  "IP",
  "CPI",
  "FFR"
)


# Reduced-form IRFs with bootstrap confidence intervals


irf_results <- irf(
  fit_aic,
  impulse = variables,
  response = variables,
  n.ahead = irf_horizon,
  ortho = FALSE,
  boot = TRUE,
  ci = irf_ci,
  runs = irf_runs,
  seed = 123
)


# Extract IRFs by impulse

irf_ip <- irf_results$irf$IP
irf_cpi <- irf_results$irf$CPI
irf_ffr <- irf_results$irf$FFR

lower_ip <- irf_results$Lower$IP
lower_cpi <- irf_results$Lower$CPI
lower_ffr <- irf_results$Lower$FFR

upper_ip <- irf_results$Upper$IP
upper_cpi <- irf_results$Upper$CPI
upper_ffr <- irf_results$Upper$FFR


# Significance of responses to a policy-rate impulse

#ffr_significance <- data.frame(
#  Horizon = 0:irf_horizon,

#  IP = (
#    lower_ffr[, "IP"] > 0 |
#    upper_ffr[, "IP"] < 0
#  ),

#  CPI = (
#    lower_ffr[, "CPI"] > 0 |
#    upper_ffr[, "CPI"] < 0
#  ),

#  FFR = (
#    lower_ffr[, "FFR"] > 0 |
#    upper_ffr[, "FFR"] < 0
#  )
#)


# Recover point-estimate responses in levels

# Variables in the estimated VAR:

# IP  = Δ log(INDPRO)       -> accumulate once
# CPI = Δ² log(CPIAUCSL)    -> accumulate twice
# FFR = Δ FEDFUNDS          -> accumulate once

to_levels <- function(irf_matrix) {

  data.frame(
    Horizon = 0:(nrow(irf_matrix) - 1),

    IP = cumsum(
      irf_matrix[, "IP"]
    ),

    CPI = cumsum(
      cumsum(
        irf_matrix[, "CPI"]
      )
    ),

    FFR = cumsum(
      irf_matrix[, "FFR"]
    )
  )
}


# Responses in levels for each impulse

irf_ip_levels <- to_levels(
  irf_ip
)

irf_cpi_levels <- to_levels(
  irf_cpi
)

irf_ffr_levels <- to_levels(
  irf_ffr
)

```

# Impulse Response Functions
## Impulse Response Functions {.t}

\vspace{2em}

\normalsize

\textcolor{navyblue}{$\bullet$} An impulse response function traces how one variable reacts over time to an innovation in another variable

\vspace{0.35em}

\small

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  Impulse: variable receiving the initial innovation

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  Response: variable whose reaction is followed over subsequent horizons

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  The IRF summarizes the sign, magnitude and persistence of the response

\end{itemize}

---

## IRF Implementation {.t}

\vspace{2em}

\normalsize

\textcolor{navyblue}{$\bullet$} Compute IRFs using \texttt{irf()} from the \texttt{vars} package

\vspace{0.4em}

\footnotesize

\texttt{irf(fit\_aic,}

\texttt{\hspace{1em} impulse = variables,}

\texttt{\hspace{1em} response = variables,}

\texttt{\hspace{1em} n.ahead = 48,}

\texttt{\hspace{1em} ortho = FALSE,}

\texttt{\hspace{1em} boot = TRUE)}

---

## IRF Point Estimates{.t}

\vspace{-1.8em}

```{r}
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.0
#| fig-align: center

impulses <- c("IP", "CPI", "FFR")
responses <- c("IP", "CPI", "FFR")

par(
  mfrow = c(3, 3),
  mar = c(1.3, 2.2, 1.2, 0.4),
  oma = c(2.2, 4.2, 2.0, 0.2),
  cex = 0.8
)

for (i in seq_along(impulses)) {

  impulse <- impulses[i]

  for (j in seq_along(responses)) {

    response <- responses[j]

    y <- irf_results$irf[[impulse]][, response]
    horizon <- 0:(length(y) - 1)

    plot(
      horizon,
      y,
      type = "l",
      ylim = range(c(y, 0)),
      xlab = "",
      ylab = "",
      main = "",
      axes = FALSE,
      lwd = 1.2
    )

    abline(
      h = 0,
      lty = 2,
      col = "grey40"
    )

    axis(
    1,
    at = seq(
      0,
      irf_horizon,
      by = 5
    ),
    cex.axis = 0.55
  )

  axis(
    2,
    cex.axis = 0.55
  )
  
    # Response labels
    if (i == 1) {
      mtext(
        response,
        side = 3,
        line = 0.15,
        font = 2,
        cex = 0.7
      )
    }
  }
}

# Impulse labels
mtext(
  "IP",
  side = 2,
  outer = TRUE,
  at = 5 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "CPI",
  side = 2,
  outer = TRUE,
  at = 3 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "FFR",
  side = 2,
  outer = TRUE,
  at = 1 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

# Common labels
mtext(
  "Response",
  side = 3,
  outer = TRUE,
  line = 0.9,
  font = 2,
  cex = 0.75
)

mtext(
  "Impulse",
  side = 2,
  outer = TRUE,
  line = 2.0,
  font = 2,
  cex = 0.75
)

mtext(
  "Horizon (months)",
  side = 1,
  outer = TRUE,
  line = 1.0,
  cex = 0.75
)

par(mfrow = c(1, 1))
```

------------------------------------------------------------------------

## Responses to an IP Innovation {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Impulse: IP, $\Delta\log(INDPRO)$

\vspace{0.4em}

\small

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  IP — strongly positive on impact, followed by rapid decay and a small negative overshoot

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  CPI — alternating positive and negative movements that become progressively smaller

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  FFR — large positive short-run response followed by oscillations around zero

\end{itemize}

\vspace{0.4em}

\normalsize

\textcolor{navyblue}{$\bullet$} Most of the effect is concentrated at short horizons and eventually dies out

------------------------------------------------------------------------

## Responses to a CPI Innovation {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Impulse: CPI, $\Delta^2\log(CPI)$

\vspace{0.4em}

\small

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  IP — initially positive, followed by a negative overshoot and gradual convergence toward zero

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  CPI — positive on impact, followed by a sharp negative correction and mean reversion

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  FFR — large short-run fluctuations that diminish over time

\end{itemize}

\vspace{0.4em}

\normalsize

\textcolor{navyblue}{$\bullet$} The responses are again concentrated mainly in the first 15--20 months

------------------------------------------------------------------------

## Responses to an FFR Innovation{.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Impulse: FFR, $\Delta FFR$

\vspace{0.4em}

\small

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  IP — becomes negative after a few months and gradually returns toward zero

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  CPI — oscillates around zero with declining amplitude

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  FFR — positive on impact, followed by a reversal and damped oscillations

\end{itemize}



------------------------------------------------------------------------

## The Price Puzzle {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Bernanke et al. (2005) discuss the price puzzle

\vspace{0.4em}

\small

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  A contractionary monetary-policy innovation (represented by an increase in the FFR) is followed by an increase in the price level

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  The relevant relationship in our VAR is an FFR impulse and a CPI response

\end{itemize}

\vspace{0.5em}

\normalsize

\textcolor{navyblue}{$\bullet$} CPI enters the VAR as $\Delta^2\log(CPI)$

\vspace{0.3em}

\small

\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}] The transformed CPI response therefore cannot by itself determine whether the price level rises

\end{itemize}

------------------------------------------------------------------------

## Bootstrap Confidence Intervals{.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Point estimates do not show the sampling uncertainty surrounding the IRFs

\vspace{0.4em}

\textcolor{navyblue}{$\bullet$} With \texttt{boot = TRUE}, \texttt{irf()} uses a residual bootstrap

\vspace{0.35em}

\footnotesize

\begin{enumerate}

  \item Estimate the VAR and obtain its residuals

  \item Resample centered residuals with replacement

  \item Generate a bootstrap time series recursively from the estimated VAR

  \item Re-estimate the VAR using the bootstrap sample

  \item Recompute the IRFs

  \item Repeat the procedure for all bootstrap replications

\end{enumerate}

------------------------------------------------------------------------

## Constructing the Confidence Interval {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} Each bootstrap replication produces a new estimate of the IRF at every horizon

\vspace{0.4em}

\textcolor{navyblue}{$\bullet$} The \texttt{vars} package uses a standard percentile interval

\vspace{0.4em}

$$CI_h^{95\%}
=
\left[
\widehat{\theta}^{*}_{h,0.025},
\widehat{\theta}^{*}_{h,0.975}
\right]$$

\vspace{0.4em}

\small
\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}] The lower and upper bounds are the 2.5th and 97.5th percentiles of the bootstrap IRF distribution at horizon $h$

\end{itemize}

\vspace{0.4em}

\normalsize

\textcolor{navyblue}{$\bullet$} Pointwise confidence intervals: uncertainty is assessed separately at each horizon

------------------------------------------------------------------------

## IRFs with 95% Bootstrap Bands{.t}

\vspace{-1.8em}

```{r}
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.0
#| fig-align: center

impulses <- c("IP", "CPI", "FFR")
responses <- c("IP", "CPI", "FFR")

par(
  mfrow = c(3, 3),
  mar = c(1.3, 2.2, 1.2, 0.4),
  oma = c(2.2, 4.2, 2.0, 0.2),
  cex = 0.8
)

for (i in seq_along(impulses)) {

  impulse <- impulses[i]

  for (j in seq_along(responses)) {

    response <- responses[j]

    point <- irf_results$irf[[impulse]][, response]
    lower <- irf_results$Lower[[impulse]][, response]
    upper <- irf_results$Upper[[impulse]][, response]

    horizon <- 0:(length(point) - 1)

    plot_range <- range(
      c(lower, upper, point, 0)
    )

    # Empty plotting region
    plot(
      horizon,
      point,
      type = "n",
      ylim = plot_range,
      xlab = "",
      ylab = "",
      main = "",
      axes = FALSE
    )

    # 95% bootstrap confidence interval
    polygon(
      c(horizon, rev(horizon)),
      c(lower, rev(upper)),
      col = adjustcolor(
        "grey70",
        alpha.f = 0.45
      ),
      border = NA
    )

    # Zero line
    abline(
      h = 0,
      lty = 3,
      col = "grey40"
    )

    # Point estimate
    lines(
      horizon,
      point,
      lwd = 1.2
    )

    axis(
  1,
  at = seq(
    0,
    irf_horizon,
    by = 5
  ),
  cex.axis = 0.55
)

axis(
  2,
  cex.axis = 0.55
)

    # Response labels
    if (i == 1) {
      mtext(
        response,
        side = 3,
        line = 0.15,
        font = 2,
        cex = 0.7
      )
    }
  }
}

# Impulse labels
mtext(
  "IP",
  side = 2,
  outer = TRUE,
  at = 5 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "CPI",
  side = 2,
  outer = TRUE,
  at = 3 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "FFR",
  side = 2,
  outer = TRUE,
  at = 1 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

# Common labels
mtext(
  "Response",
  side = 3,
  outer = TRUE,
  line = 0.9,
  font = 2,
  cex = 0.75
)

mtext(
  "Impulse",
  side = 2,
  outer = TRUE,
  line = 2.0,
  font = 2,
  cex = 0.75
)

mtext(
  "Horizon (months)",
  side = 1,
  outer = TRUE,
  line = 1.0,
  cex = 0.75
)

par(mfrow = c(1, 1))
```

------------------------------------------------------------------------

## Significance of the IRFs {.t}

\vspace{2em}

\normalsize

\textcolor{navyblue}{$\bullet$} A response is pointwise significant at the 5% level when its 95% confidence band excludes zero

\vspace{0.3em}

\small

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  The clearest evidence is concentrated in short-run own-responses

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  Several cross-responses are imprecisely estimated and their bands include zero over much of the horizon

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  Statistical significance generally weakens as the horizon increases

\end{itemize}

------------------------------------------------------------------------

## Recovering IRFs in Levels

\normalsize

\textcolor{navyblue}{$\bullet$} The VAR is estimated using stationary transformations rather than the original levels

\vspace{0.4em}

\textcolor{navyblue}{$\bullet$} A response in first differences is converted to a level response by accumulating the IRF (Lütkepohl, 2005)

\vspace{0.35em}

\footnotesize
\begin{itemize}

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}] For industrial production:

$$
IRF_h^{\log(INDPRO)}
=
\sum_{s=0}^{h}
IRF_s^{\Delta\log(INDPRO)}
$$

\item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}] For the federal funds rate:

$$
IRF_h^{FFR}
=
\sum_{s=0}^{h}
IRF_s^{\Delta FFR}
$$

\end{itemize}

------------------------------------------------------------------------

## Recovering the CPI Level Response {.t}

\vspace{1em}

\normalsize

\textcolor{navyblue}{$\bullet$} CPI enters the VAR in second log differences $\Delta^2\log(CPI_t)$

\textcolor{navyblue}{$\bullet$} Its response therefore requires two accumulations

\vspace{0.3em}

\footnotesize
\begin{itemize}

\item[] 1. Recover the inflation response:

$$
IRF_h^{\Delta\log(CPI)}
=
\sum_{s=0}^{h}
IRF_s^{\Delta^2\log(CPI)}
$$

\item[] 2. Recover the log price-level response:

$$
IRF_h^{\log(CPI)}
=
\sum_{j=0}^{h}
IRF_j^{\Delta\log(CPI)}
$$

\end{itemize}

\vspace{0.3em}

\normalsize

\textcolor{navyblue}{$\bullet$} The resulting CPI IRF describes the cumulative response of the log price level

------------------------------------------------------------------------

## IRFs Reconstructed in Levels{.t}

\vspace{-1.8em}

```{r}
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.0
#| fig-align: center

impulses <- c("IP", "CPI", "FFR")
responses <- c("IP", "CPI", "FFR")

level_irfs <- list(
  IP = irf_ip_levels,
  CPI = irf_cpi_levels,
  FFR = irf_ffr_levels
)

par(
  mfrow = c(3, 3),
  mar = c(1.3, 2.2, 1.2, 0.4),
  oma = c(2.2, 4.2, 2.0, 0.2),
  cex = 0.8
)

for (i in seq_along(impulses)) {

  impulse <- impulses[i]

  for (j in seq_along(responses)) {

    response <- responses[j]

    y <- level_irfs[[impulse]][[response]]
    horizon <- level_irfs[[impulse]]$Horizon

    plot(
      horizon,
      y,
      type = "l",
      ylim = range(c(y, 0)),
      xlab = "",
      ylab = "",
      main = "",
      axes = FALSE,
      lwd = 1.2
    )

    abline(
      h = 0,
      lty = 2,
      col = "grey40"
    )

    axis(
  1,
  at = seq(
    0,
    irf_horizon,
    by = 5
  ),
  cex.axis = 0.55
)

axis(
  2,
  cex.axis = 0.55
)

    # Response labels
    if (i == 1) {
      mtext(
        response,
        side = 3,
        line = 0.15,
        font = 2,
        cex = 0.7
      )
    }
  }
}

# Impulse labels
mtext(
  "IP",
  side = 2,
  outer = TRUE,
  at = 5 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "CPI",
  side = 2,
  outer = TRUE,
  at = 3 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "FFR",
  side = 2,
  outer = TRUE,
  at = 1 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

# Common labels
mtext(
  "Response",
  side = 3,
  outer = TRUE,
  line = 0.9,
  font = 2,
  cex = 0.75
)

mtext(
  "Impulse",
  side = 2,
  outer = TRUE,
  line = 2.0,
  font = 2,
  cex = 0.75
)

mtext(
  "Horizon (months)",
  side = 1,
  outer = TRUE,
  line = 1.0,
  cex = 0.75
)

par(mfrow = c(1, 1))
```

------------------------------------------------------------------------

## Interpretation of the Level Responses {.t}

\vspace{2em}

\normalsize

\textcolor{navyblue}{$\bullet$} Accumulation changes the interpretation of the responses

\vspace{0.4em}

\footnotesize

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  An IP innovation is followed by higher industrial production and a higher CPI level over the horizon

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  A CPI innovation generates cumulative movements across all three variables in levels

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  An FFR innovation is associated with lower industrial production and a higher CPI level

\end{itemize}

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Unlike the IRFs of the differenced variables, the reconstructed level responses need not converge back to zero

------------------------------------------------------------------------

## Interpretation of the Level Responses {.t}

\vspace{2em}

\normalsize

\textcolor{navyblue}{$\bullet$} Accumulation changes the interpretation of the responses

\vspace{0.4em}

\footnotesize

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  An IP innovation is followed by higher industrial production and a higher CPI level over the horizon

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  A CPI innovation generates cumulative movements across all three variables in levels

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  An FFR innovation is associated with lower industrial production and a higher CPI level

\end{itemize}

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Unlike the IRFs of the differenced variables, the reconstructed level responses need not converge back to zero

\vspace{0.45em}

\textcolor{navyblue}{$\Rightarrow$} \underline{So, does the price puzzle occur?}

------------------------------------------------------------------------

## Interpretation of the Level Responses {.t}

\vspace{2em}

\normalsize

\textcolor{navyblue}{$\bullet$} Accumulation changes the interpretation of the responses

\vspace{0.4em}

\footnotesize

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  An IP innovation is followed by higher industrial production and a higher CPI level over the horizon

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  A CPI innovation generates cumulative movements across all three variables in levels

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  An FFR innovation is associated with lower industrial production and a higher CPI level

\end{itemize}

\vspace{0.45em}

\normalsize

\textcolor{navyblue}{$\bullet$} Unlike the IRFs of the differenced variables, the reconstructed level responses need not converge back to zero

\vspace{0.45em}

\textcolor{navyblue}{$\Rightarrow$} \underline{So, does the price puzzle occur?} \quad \textbf{Yes and no.}

------------------------------------------------------------------------

## Reduced Form and Structural Analysis

\normalsize

\textcolor{navyblue}{$\bullet$} Reduced-form IRFs describe responses to VAR forecast errors

\vspace{0.4em}

\textcolor{navyblue}{$\bullet$} These innovations are not automatically interpretable as specific economic shocks

\vspace{0.2em}

\footnotesize

\begin{itemize}

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  Reduced-form residuals may be contemporaneously correlated

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  A policy-rate innovation need not represent an exogenous monetary-policy shock

  \item[\raise0.15ex\hbox{\textcolor{navyblue}{\tiny$\circ$}}]
  Structural analysis requires additional identifying restrictions

\end{itemize}

\vspace{0.45em}

\normalsize

The present IRFs are informative about dynamic propagation, but causal structural interpretation requires an identified structural VAR

---

## {.plain}

\vfill

\begin{center}
{\LARGE \textcolor{navyblue}{Thank you for your attention!}}
\end{center}

\vfill
