---
title: "visual2"
format: pdf
editor: visual
---

### 0. Settings

```{r}
<<<<<<< HEAD
# install.packages(c("bootUR", "vars"))
# install.packages("lpirfs")
=======
# install.packages(c("bootUR", "vars", "car"))

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
library(bootUR)
library(vars)
library(car)
library(lpirfs)
```

### 1. Data preparation

```{r}
# Load the FRED-MD CSV
raw <- read.csv("v2-2026-rev-08-MD.csv", check.names = FALSE)

# FRED-MD transformation codes for the selected variables
fred_codes <- unlist(
  raw[1, c("INDPRO", "CPIAUCSL", "FEDFUNDS")]
)

# Keep only the variables used in the case study
dat <- raw[, c(
  "sasdate",
  "INDPRO",
  "CPIAUCSL",
  "FEDFUNDS"
)]

names(dat)[1] <- "Date"

dat$Date <- as.Date(
  dat$Date,
  format = "%m/%d/%Y"
)

dat <- dat[
  !is.na(dat$Date) &
    dat$Date <= as.Date("2019-12-31"),
]

dat <- dat[order(dat$Date), ]

fred_codes
```

### 2. Series transformation

```{r}
dat_transformed <- data.frame(
  Date = dat$Date,

  INDPRO = c(
    NA,
    diff(log(dat$INDPRO))
  ),

  CPIAUCSL = c(
    NA,
    NA,
    diff(
      log(dat$CPIAUCSL),
      differences = 2
    )
  ),

  FEDFUNDS = c(
    NA,
    diff(dat$FEDFUNDS)
  )
)

# Common sample starts in March 1959
dat_common <- dat[-c(1, 2), ]

dat_transformed <- dat_transformed[
  -c(1, 2),
]

common_start <- c(
  as.integer(
    format(dat_common$Date[1], "%Y")
  ),
  as.integer(
    format(dat_common$Date[1], "%m")
  )
)

# Original series
levels_ts <- ts(
  dat_common[, c(
    "INDPRO",
    "CPIAUCSL",
    "FEDFUNDS"
  )],
  start = common_start,
  frequency = 12
)

# Transformed series
transformed_ts <- ts(
  dat_transformed[, c(
    "INDPRO",
    "CPIAUCSL",
    "FEDFUNDS"
  )],
  start = common_start,
  frequency = 12
)

range(dat_transformed$Date)
```

### 3. Series in levels

```{r}
plot(
  levels_ts[, "INDPRO"],
  main = "Industrial Production",
  xlab = "Year",
  ylab = "Index"
)
```

```{r}
plot(
  levels_ts[, "CPIAUCSL"],
  main = "Consumer Price Index",
  xlab = "Year",
  ylab = "Index"
)
```

```{r}
plot(
  levels_ts[, "FEDFUNDS"],
  main = "Federal Funds Rate",
  xlab = "Year",
  ylab = "Percent"
)
```

### 4. Transformed series

```{r}
plot(
  transformed_ts[, "INDPRO"],
  main = "Output Growth",
  xlab = "Year",
  ylab = expression(
    Delta~log(INDPRO)
  )
)
```

```{r}
plot(
  transformed_ts[, "CPIAUCSL"],
  main = "Change in Inflation",
  xlab = "Year",
  ylab = expression(
    Delta^2~log(CPI)
  )
)
```

```{r}
plot(
  transformed_ts[, "FEDFUNDS"],
  main = "Change in Policy Rate",
  xlab = "Year",
  ylab = expression(
    Delta~FEDFUNDS
  )
)
```

### 5. Unit root testing

```{r}
# Series considered for the VAR

IP  <- log(dat$INDPRO)
CPI <- log(dat$CPIAUCSL)
FFR <- dat$FEDFUNDS

# Bootstrap settings

set.seed(123)
B <- 4999


# Trend specification

IP_trend <- order_integration(
  IP,
  method = "boot_adf",
  deterministics = "trend",
  B = B
)

CPI_trend <- order_integration(
  CPI,
  method = "boot_adf",
  deterministics = "trend",
  B = B
)

FFR_trend <- order_integration(
  FFR,
  method = "boot_adf",
  deterministics = "trend",
  B = B
)

trend_orders <- c(
  IP  = unname(IP_trend$order_int),
  CPI = unname(CPI_trend$order_int),
  FFR = unname(FFR_trend$order_int)
)


# Intercept specification

IP_intercept <- order_integration(
  IP,
  method = "boot_adf",
  deterministics = "intercept",
  B = B
)

CPI_intercept <- order_integration(
  CPI,
  method = "boot_adf",
  deterministics = "intercept",
  B = B
)

FFR_intercept <- order_integration(
  FFR,
  method = "boot_adf",
  deterministics = "intercept",
  B = B
)

intercept_orders <- c(
  IP  = unname(IP_intercept$order_int),
  CPI = unname(CPI_intercept$order_int),
  FFR = unname(FFR_intercept$order_int)
)


# Compare deterministic specifications

unit_root_table <- data.frame(
  Series = c(
    "IP",
    "CPI",
    "FFR"
  ),

  Trend = as.integer(
    trend_orders
  ),

  Intercept = as.integer(
    intercept_orders
  ),

  Different = (
    trend_orders != intercept_orders
  )
)

unit_root_table


# Manual unit roots with boot_adf()

urip1 <- boot_adf(diff(diff(IP)), deterministics = "intercept", B = B)
urip2 <- boot_adf(diff(IP), deterministics = "intercept", B = B)
urip3 <- boot_adf(IP, deterministics = "trend", B = B)

urcpi1 <- boot_adf(diff(diff(CPI)), deterministics = "intercept", B = B)
urcpi2 <- boot_adf(diff(CPI), deterministics = "intercept", B = B)
urcpi3 <- boot_adf(CPI, deterministics = "trend", B = B)

urffr1 <- boot_adf(diff(diff(FFR)), deterministics = "intercept", B = B)
urffr2 <- boot_adf(diff(FFR), deterministics = "intercept", B = B)
urffr3 <- boot_adf(FFR, deterministics = "trend", B = B)

unit_root_table_manual <- data.frame(
  Series = c(
    "IP",
    "CPI",
    "FFR"
  ),
  
  Delta_2_P = c(
    urip1$p.value,
    urcpi1$p.value,
    urffr1$p.value
  ),
  
  Delta_1_P = c(
    urip2$p.value,
    urcpi2$p.value,
    urffr2$p.value
  ),
  
  Delta_0_P = c(
    urip3$p.value,
    urcpi3$p.value,
    urffr3$p.value
  ),
  
  Order = c(
    as.numeric(1),
    as.numeric(1),
    as.numeric(1)
  )
)

unit_root_table_manual
```

### 6. Selected integration order

```{r}
# Selected deterministic specifications:
# IP  -> trend
# CPI -> trend
# FFR -> intercept

selected_integration_table <- data.frame(
  Series = c(
    "IP",
    "CPI",
    "FFR"
  ),

  Deterministic = c(
    "Trend",
    "Trend",
    "Intercept"
  ),

  Integration_order = c(
    trend_orders["IP"],
    trend_orders["CPI"],
    intercept_orders["FFR"]
  )
)

selected_integration_table
```

### 7. VAR setup

```{r}
alpha <- 0.05

Y <- data.frame(
  Date = dat_transformed$Date,
  IP   = dat_transformed$INDPRO,
  CPI  = dat_transformed$CPIAUCSL,
  FFR  = dat_transformed$FEDFUNDS
)

x <- as.matrix(
  Y[, c(
    "IP",
    "CPI",
    "FFR"
  )]
)

stopifnot(
  all(is.finite(x))
)

c(
  observations = nrow(x),
  variables = ncol(x)
)
```

### 8. Lag order selection

```{r}
# Allow up to two years of monthly lags
lag_max <- 24

lag_selection <- VARselect(
  x,
  lag.max = lag_max,
  type = "const"
)

lag_table <- data.frame(
  Criterion = c(
    "AIC",
    "HQ",
    "SC / BIC",
    "FPE"
  ),

  Selected_lag = as.integer(c(
    lag_selection$selection[["AIC(n)"]],
    lag_selection$selection[["HQ(n)"]],
    lag_selection$selection[["SC(n)"]],
    lag_selection$selection[["FPE(n)"]]
  ))
)

lag_criteria <- lag_selection$criteria

p_aic <- as.integer(
  lag_selection$selection[["AIC(n)"]]
)

p_bic <- as.integer(
  lag_selection$selection[["SC(n)"]]
)

# Display selected lag orders
lag_table
```

```{r}
# Display criterion values for all candidate lags
lag_criteria
```

### 9. Estimation VAR models

```{r}

fit_aic <- VAR(
  x,
  p = p_aic,
  type = "const"
)

fit_bic <- VAR(
  x,
  p = p_bic,
  type = "const"
)

c(
  AIC_selected_lag = p_aic,
  BIC_selected_lag = p_bic
  
)
```

### 10. VAR stability

```{r}

roots_aic <- roots(
  fit_aic,
  modulus = TRUE
)

roots_bic <- roots(
  fit_bic,
  modulus = TRUE
)

stability_table <- data.frame(
  Specification = c(
    paste0(
      "AIC VAR(",
      p_aic,
      ")"
    ),
    paste0(
      "SC/BIC VAR(",
      p_bic,
      ")"
    )
  ),

  Max_root = c(
    max(roots_aic),
    max(roots_bic)
  ),

  Stable = c(
    all(roots_aic < 1),
    all(roots_bic < 1)
  )
)

stability_table
```

### 11. Residual diagnostics

```{r}
# Common two-year residual-autocorrelation horizon
serial_horizon <- 24

stopifnot(
  p_aic < serial_horizon,
  p_bic < serial_horizon
)

run_var_diagnostics <- function(
  fit,
  serial_horizon
) {

  root_moduli <- roots(
    fit,
    modulus = TRUE
  )

  serial_result <- serial.test(
    fit,
    lags.pt = serial_horizon,
    type = "PT.adjusted"
  )

  arch_result <- arch.test(
    fit
  )

  normality_result <- normality.test(
    fit
  )

  list(
    lag = fit$p,

    roots = root_moduli,

    stable = all(
      root_moduli < 1
    ),

    max_root = max(
      root_moduli
    ),

    serial_test = serial_result,

    serial_p = as.numeric(
      serial_result$serial$p.value
    ),

    arch_test = arch_result,

    arch_p = as.numeric(
      arch_result$arch.mul$p.value
    ),

    normality_test = normality_result,

    normality_p = as.numeric(
      normality_result$jb.mul$JB$p.value
    )
  )
}

diag_aic <- run_var_diagnostics(
  fit_aic,
  serial_horizon
)

diag_bic <- run_var_diagnostics(
  fit_bic,
  serial_horizon
)
```

### 12. Diagnostic results

```{r}
diagnostic_table <- data.frame(
  Criterion = c(
    "AIC",
    "SC / BIC"
  ),

  Selected_lag = c(
    diag_aic$lag,
    diag_bic$lag
  ),

  Stable = c(
    diag_aic$stable,
    diag_bic$stable
  ),

  Max_root = c(
    diag_aic$max_root,
    diag_bic$max_root
  ),

  Serial_p = c(
    diag_aic$serial_p,
    diag_bic$serial_p
  ),

  ARCH_p = c(
    diag_aic$arch_p,
    diag_bic$arch_p
  ),

  Normality_p = c(
    diag_aic$normality_p,
    diag_bic$normality_p
  )
)

diagnostic_table$Serial_correlation <- ifelse(
  diagnostic_table$Serial_p < alpha,
  "Rejected",
  "Not rejected"
)

diagnostic_table$ARCH_effects <- ifelse(
  diagnostic_table$ARCH_p < alpha,
  "Rejected",
  "Not rejected"
)

diagnostic_table$Normality <- ifelse(
  diagnostic_table$Normality_p < alpha,
  "Rejected",
  "Not rejected"
)

diagnostic_table
```

```{r}
# Serial correlation
diag_aic$serial_test
diag_bic$serial_test
```

```{r}
# ARCH effects
diag_aic$arch_test
diag_bic$arch_test
```

```{r}
# Residual normality
diag_aic$normality_test
diag_bic$normality_test
```

<<<<<<< HEAD
### 13.  Granger

```{r}
#install.packages("car")
library(car)

# Granger causality

#alpha <- 0.05


=======
### 13. Granger causality

```{r}

# Granger causality

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
# CPI and FFR -> IP
gc_IP <- causality(
  fit_aic,
  cause = c("CPI", "FFR")
)$Granger

# IP and FFR -> CPI
gc_CPI <- causality(
  fit_aic,
  cause = c("IP", "FFR")
)$Granger

# IP and CPI -> FFR
gc_FFR <- causality(
  fit_aic,
  cause = c("IP", "CPI")
)$Granger


# Collect the three block-test results
granger_block_table <- data.frame(
  x = c(
    "CPI + FFR",
    "IP + FFR",
    "IP + CPI"
  ),

  y = c(
    "IP",
    "CPI",
    "FFR"
  ),

  F_statistic = c(
    as.numeric(gc_IP$statistic),
    as.numeric(gc_CPI$statistic),
    as.numeric(gc_FFR$statistic)
  ),

  p_value = c(
    gc_IP$p.value,
    gc_CPI$p.value,
    gc_FFR$p.value
  )
)

granger_block_table$Decision <- ifelse(
  granger_block_table$p_value < alpha,
  "Reject H0",
  "Do not reject H0"
)

<<<<<<< HEAD


=======
```

### 14. Conditional pairwise Granger causality

```{r}
>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
# Six (conditional) pairwise tests


# IP -> CPI
#
# H0:
# IP.l1 = IP.l2 = ... = IP.l14 = 0
# in the CPI equation


test_IP_CPI <- linearHypothesis(
  fit_aic$varresult[["CPI"]],
  paste0(
    "IP.l",
    1:p_aic,
    " = 0"
  )
)


# IP -> FFR

test_IP_FFR <- linearHypothesis(
  fit_aic$varresult[["FFR"]],
  paste0(
    "IP.l",
    1:p_aic,
    " = 0"
  )
)


# CPI -> IP

test_CPI_IP <- linearHypothesis(
  fit_aic$varresult[["IP"]],
  paste0(
    "CPI.l",
    1:p_aic,
    " = 0"
  )
)


# CPI -> FFR

test_CPI_FFR <- linearHypothesis(
  fit_aic$varresult[["FFR"]],
  paste0(
    "CPI.l",
    1:p_aic,
    " = 0"
  )
)



# FFR -> IP

test_FFR_IP <- linearHypothesis(
  fit_aic$varresult[["IP"]],
  paste0(
    "FFR.l",
    1:p_aic,
    " = 0"
  )
)



# FFR -> CPI

test_FFR_CPI <- linearHypothesis(
  fit_aic$varresult[["CPI"]],
  paste0(
    "FFR.l",
    1:p_aic,
    " = 0"
  )
)


# Collect results

granger_pairwise_table <- data.frame(
  x = c(
    "IP",
    "IP",
    "CPI",
    "CPI",
    "FFR",
    "FFR"
  ),

  y = c(
    "CPI",
    "FFR",
    "IP",
    "FFR",
    "IP",
    "CPI"
  ),

  F_statistic = c(
    test_IP_CPI$F[2],
    test_IP_FFR$F[2],
    test_CPI_IP$F[2],
    test_CPI_FFR$F[2],
    test_FFR_IP$F[2],
    test_FFR_CPI$F[2]
  ),

  p_value = c(
    test_IP_CPI$`Pr(>F)`[2],
    test_IP_FFR$`Pr(>F)`[2],
    test_CPI_IP$`Pr(>F)`[2],
    test_CPI_FFR$`Pr(>F)`[2],
    test_FFR_IP$`Pr(>F)`[2],
    test_FFR_CPI$`Pr(>F)`[2]
  )
)


# Decision at the 5% level
granger_pairwise_table$Decision <- ifelse(
  granger_pairwise_table$p_value < alpha,
  "Reject H0",
  "Do not reject H0"
)

<<<<<<< HEAD

=======
```

### 15. Granger tables

```{r}
>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f

# Tables

granger_block_slide <- granger_block_table

granger_block_slide$F_statistic <- round(
  granger_block_slide$F_statistic,
  2
)

granger_block_slide$p_value <- format.pval(
  granger_block_slide$p_value,
  digits = 3,
  eps = 0.001
)


granger_pairwise_slide <- granger_pairwise_table

granger_pairwise_slide$F_statistic <- round(
  granger_pairwise_slide$F_statistic,
  2
)

granger_pairwise_slide$p_value <- format.pval(
  granger_pairwise_slide$p_value,
  digits = 3,
  eps = 0.001
)

```

<<<<<<< HEAD
```{r}
#| echo: false
#| results: asis

knitr::kable(
  granger_block_slide,
  row.names = FALSE,
  align = "c"
)
```

Pairwise:

```{r}
#| echo: false
#| results: asis

knitr::kable(
  granger_pairwise_slide,
  row.names = FALSE,
  align = "c"
)


```

### 14.  IRFs in reduced VAR form

```{r}
#| include: false
=======

### 16. Impulse response functions

```{r}
>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f

# IRF settings

irf_horizon <- 48
<<<<<<< HEAD
irf_runs <- 1000
=======
irf_runs <- 4999
>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
irf_ci <- 0.95

variables <- c(
  "IP",
  "CPI",
  "FFR"
)


# Reduced-form IRFs with bootstrap confidence intervals


irf_results <- irf(
  fit_aic,
  impulse = variables,
  response = variables,
  n.ahead = irf_horizon,
  ortho = FALSE,
  boot = TRUE,
  ci = irf_ci,
  runs = irf_runs,
  seed = 123
)


# Extract IRFs by impulse

irf_ip <- irf_results$irf$IP
irf_cpi <- irf_results$irf$CPI
irf_ffr <- irf_results$irf$FFR

lower_ip <- irf_results$Lower$IP
lower_cpi <- irf_results$Lower$CPI
lower_ffr <- irf_results$Lower$FFR

upper_ip <- irf_results$Upper$IP
upper_cpi <- irf_results$Upper$CPI
upper_ffr <- irf_results$Upper$FFR


<<<<<<< HEAD
# Significance of responses to a policy-rate impulse

ffr_significance <- data.frame(
  Horizon = 0:irf_horizon,
  
  IP = (
    lower_ffr[, "IP"] > 0 |
      upper_ffr[, "IP"] < 0
  ),
  
  CPI = (
    lower_ffr[, "CPI"] > 0 |
      upper_ffr[, "CPI"] < 0
  ),
  
  FFR = (
    lower_ffr[, "FFR"] > 0 |
      upper_ffr[, "FFR"] < 0
  )
)

=======
```

### 17. Recovering IRFs in levels

```{r}
>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f

# Recover point-estimate responses in levels

# Variables in the estimated VAR:

<<<<<<< HEAD
# IP  = Δ log(INDPRO)       -> accumulate once
# CPI = Δ² log(CPIAUCSL)    -> accumulate twice
# FFR = Δ FEDFUNDS          -> accumulate once

to_levels <- function(irf_matrix) {
  
  data.frame(
    Horizon = 0:(nrow(irf_matrix) - 1),
    
    IP = cumsum(
      irf_matrix[, "IP"]
    ),
    
=======
# IP  = difflog(INDPRO)       -> accumulate once
# CPI = diff2log(CPIAUCSL)    -> accumulate twice
# FFR = diffFEDFUNDS          -> accumulate once

to_levels <- function(irf_matrix) {

  data.frame(
    Horizon = 0:(nrow(irf_matrix) - 1),

    IP = cumsum(
      irf_matrix[, "IP"]
    ),

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    CPI = cumsum(
      cumsum(
        irf_matrix[, "CPI"]
      )
    ),
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    FFR = cumsum(
      irf_matrix[, "FFR"]
    )
  )
}


# Responses in levels for each impulse

irf_ip_levels <- to_levels(
  irf_ip
)

irf_cpi_levels <- to_levels(
  irf_cpi
)

irf_ffr_levels <- to_levels(
  irf_ffr
)

```

<<<<<<< HEAD
=======
### 18. IRF points estimates

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
```{r}

impulses <- c("IP", "CPI", "FFR")
responses <- c("IP", "CPI", "FFR")

par(
  mfrow = c(3, 3),
  mar = c(1.3, 2.2, 1.2, 0.4),
  oma = c(2.2, 4.2, 2.0, 0.2),
  cex = 0.8
)

for (i in seq_along(impulses)) {
<<<<<<< HEAD
  
  impulse <- impulses[i]
  
  for (j in seq_along(responses)) {
    
    response <- responses[j]
    
    y <- irf_results$irf[[impulse]][, response]
    horizon <- 0:(length(y) - 1)
    
=======

  impulse <- impulses[i]

  for (j in seq_along(responses)) {

    response <- responses[j]

    y <- irf_results$irf[[impulse]][, response]
    horizon <- 0:(length(y) - 1)

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    plot(
      horizon,
      y,
      type = "l",
      ylim = range(c(y, 0)),
      xlab = "",
      ylab = "",
      main = "",
      axes = FALSE,
      lwd = 1.2
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    abline(
      h = 0,
      lty = 2,
      col = "grey40"
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    axis(
      1,
      cex.axis = 0.55
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    axis(
      2,
      cex.axis = 0.55
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    # Response labels
    if (i == 1) {
      mtext(
        response,
        side = 3,
        line = 0.15,
        font = 2,
        cex = 0.7
      )
    }
  }
}

# Impulse labels
mtext(
  "IP",
  side = 2,
  outer = TRUE,
  at = 5 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "CPI",
  side = 2,
  outer = TRUE,
  at = 3 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "FFR",
  side = 2,
  outer = TRUE,
  at = 1 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

# Common labels
mtext(
  "Response",
  side = 3,
  outer = TRUE,
  line = 0.9,
  font = 2,
  cex = 0.75
)

mtext(
  "Impulse",
  side = 2,
  outer = TRUE,
  line = 2.0,
  font = 2,
  cex = 0.75
)

mtext(
  "Horizon (months)",
  side = 1,
  outer = TRUE,
  line = 1.0,
  cex = 0.75
)

par(mfrow = c(1, 1))
<<<<<<< HEAD
```

With confidence bands:

```{r}
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.0
#| fig-align: center
=======

```

### 19. IRFs with bootstrap CBs

```{r}
>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f

impulses <- c("IP", "CPI", "FFR")
responses <- c("IP", "CPI", "FFR")

par(
  mfrow = c(3, 3),
  mar = c(1.3, 2.2, 1.2, 0.4),
  oma = c(2.2, 4.2, 2.0, 0.2),
  cex = 0.8
)

for (i in seq_along(impulses)) {
<<<<<<< HEAD
  
  impulse <- impulses[i]
  
  for (j in seq_along(responses)) {
    
    response <- responses[j]
    
    point <- irf_results$irf[[impulse]][, response]
    lower <- irf_results$Lower[[impulse]][, response]
    upper <- irf_results$Upper[[impulse]][, response]
    
    horizon <- 0:(length(point) - 1)
    
    plot_range <- range(
      c(lower, upper, point, 0)
    )
    
=======

  impulse <- impulses[i]

  for (j in seq_along(responses)) {

    response <- responses[j]

    point <- irf_results$irf[[impulse]][, response]
    lower <- irf_results$Lower[[impulse]][, response]
    upper <- irf_results$Upper[[impulse]][, response]

    horizon <- 0:(length(point) - 1)

    plot_range <- range(
      c(lower, upper, point, 0)
    )

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    # Empty plotting region
    plot(
      horizon,
      point,
      type = "n",
      ylim = plot_range,
      xlab = "",
      ylab = "",
      main = "",
      axes = FALSE
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    # 95% bootstrap confidence interval
    polygon(
      c(horizon, rev(horizon)),
      c(lower, rev(upper)),
      col = adjustcolor(
        "grey70",
        alpha.f = 0.45
      ),
      border = NA
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    # Zero line
    abline(
      h = 0,
      lty = 3,
      col = "grey40"
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    # Point estimate
    lines(
      horizon,
      point,
      lwd = 1.2
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    axis(
      1,
      cex.axis = 0.55
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    axis(
      2,
      cex.axis = 0.55
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    # Response labels
    if (i == 1) {
      mtext(
        response,
        side = 3,
        line = 0.15,
        font = 2,
        cex = 0.7
      )
    }
  }
}

# Impulse labels
mtext(
  "IP",
  side = 2,
  outer = TRUE,
  at = 5 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "CPI",
  side = 2,
  outer = TRUE,
  at = 3 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "FFR",
  side = 2,
  outer = TRUE,
  at = 1 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

# Common labels
mtext(
  "Response",
  side = 3,
  outer = TRUE,
  line = 0.9,
  font = 2,
  cex = 0.75
)

mtext(
  "Impulse",
  side = 2,
  outer = TRUE,
  line = 2.0,
  font = 2,
  cex = 0.75
)

mtext(
  "Horizon (months)",
  side = 1,
  outer = TRUE,
  line = 1.0,
  cex = 0.75
)

par(mfrow = c(1, 1))
<<<<<<< HEAD
```

Back to levels:

```{r}
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.0
#| fig-align: center
=======

```


### 20. IRFs reconstructed in levels

```{r}
>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f

impulses <- c("IP", "CPI", "FFR")
responses <- c("IP", "CPI", "FFR")

<<<<<<< HEAD
level_irfs <- list(
  IP = irf_ip_levels,
  CPI = irf_cpi_levels,
  FFR = irf_ffr_levels
)

=======
>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
par(
  mfrow = c(3, 3),
  mar = c(1.3, 2.2, 1.2, 0.4),
  oma = c(2.2, 4.2, 2.0, 0.2),
  cex = 0.8
)

for (i in seq_along(impulses)) {
<<<<<<< HEAD
  
  impulse <- impulses[i]
  
  for (j in seq_along(responses)) {
    
    response <- responses[j]
    
    y <- level_irfs[[impulse]][[response]]
    horizon <- level_irfs[[impulse]]$Horizon
    
    plot(
      horizon,
      y,
      type = "l",
      ylim = range(c(y, 0)),
      xlab = "",
      ylab = "",
      main = "",
      axes = FALSE,
      lwd = 1.2
    )
    
    abline(
      h = 0,
      lty = 2,
      col = "grey40"
    )
    
=======

  impulse <- impulses[i]

  for (j in seq_along(responses)) {

    response <- responses[j]

    point <- irf_results$irf[[impulse]][, response]
    lower <- irf_results$Lower[[impulse]][, response]
    upper <- irf_results$Upper[[impulse]][, response]

    horizon <- 0:(length(point) - 1)

    plot_range <- range(
      c(lower, upper, point, 0)
    )

    # Empty plotting region
    plot(
      horizon,
      point,
      type = "n",
      ylim = plot_range,
      xlab = "",
      ylab = "",
      main = "",
      axes = FALSE
    )

    # 95% bootstrap confidence interval
    polygon(
      c(horizon, rev(horizon)),
      c(lower, rev(upper)),
      col = adjustcolor(
        "grey70",
        alpha.f = 0.45
      ),
      border = NA
    )

    # Zero line
    abline(
      h = 0,
      lty = 3,
      col = "grey40"
    )

    # Point estimate
    lines(
      horizon,
      point,
      lwd = 1.2
    )

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    axis(
      1,
      cex.axis = 0.55
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    axis(
      2,
      cex.axis = 0.55
    )
<<<<<<< HEAD
    
=======

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
    # Response labels
    if (i == 1) {
      mtext(
        response,
        side = 3,
        line = 0.15,
        font = 2,
        cex = 0.7
      )
    }
  }
}

# Impulse labels
mtext(
  "IP",
  side = 2,
  outer = TRUE,
  at = 5 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "CPI",
  side = 2,
  outer = TRUE,
  at = 3 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

mtext(
  "FFR",
  side = 2,
  outer = TRUE,
  at = 1 / 6,
  line = 0.5,
  font = 2,
  cex = 0.7
)

# Common labels
mtext(
  "Response",
  side = 3,
  outer = TRUE,
  line = 0.9,
  font = 2,
  cex = 0.75
)

mtext(
  "Impulse",
  side = 2,
  outer = TRUE,
  line = 2.0,
  font = 2,
  cex = 0.75
)

mtext(
  "Horizon (months)",
  side = 1,
  outer = TRUE,
  line = 1.0,
  cex = 0.75
)

par(mfrow = c(1, 1))
<<<<<<< HEAD
```

### 15. Local Projections in reduced VAR form

Will use implementation made by Adammer (2019) based on Jorda's (2005) local projections paper.

```{r}
# using our fitted var fit_aic, we check the linear IRFs via LP
# make transformed data series into dataframe for the function
set.seed(333)
df_level <- data.frame(IP,CPI,FFR)
df_diff <- data.frame(IP = dat_transformed$INDPRO , CPI = dat_transformed$CPIAUCSL, FFR = dat_transformed$FEDFUNDS)

lp_level <- lp_lin(
  endog_data = df_level,
  lags_endog_lin = p_aic, 
  trend = 0, 
  shock_type = 0,
  confint = 1.96,
  hor = irf_horizon,
  use_nw = TRUE
)

lp_trans <- lp_lin(
  endog_data = df_diff,
  lags_endog_lin = p_aic, 
  trend = 0, 
  shock_type = 0,
  confint = 1.96,
  hor = irf_horizon,
  use_nw = TRUE
)


irf_orth <- irf(fit_aic, n.ahead = irf_horizon, ortho = TRUE,
                boot = TRUE, ci = 0.95, runs = 1000)

plot(lp_level)
plot(lp_trans)
par(mfrow=c(3,3))
plot(irf_orth)
summary(lp_ip)
```
=======

```

>>>>>>> efe377476f3675cdfeb54872c8e8029a9da3160f
